---
title: "More Data Transformation"
subtitle: "Lecture 5"
date: "2025-05-19"
format: 
  revealjs:
    output-file: 05-data-transform-2-slides.html
auto-stretch: false
---

## Announcements/Reminders {.smaller}

-   Lab is due TONIGHT at 11:59PM

-   Come to office hours and/or post on Ed for help!

-   Office hours are tonight from **5:30-7:30PM** in **Old Chem 203**.

-   AEs need to be pushed by the **END OF CLASS (10:45AM)!!**

-   Make sure you are in the right repository.

## Lab: Narrative {.smaller}

-   ***Boxplots, histograms, density plots:***

    -   Center: Give an idea of a 'typical value' (median, most common range, etc.);

    -   Spread: Does the data have a lot of variability?
        Or just a little?
        (IQR, outliers)

    -   Shape: Is there skew?
        Which direction?
        Is it unimodal or multimodal?

## Lab: Narrative {.smaller}

-   ***Scatter Plots***

    -   Is there a relationship between the two variables?
        Is it positive or negative?

    -   If there is a relationship, does it seem to be strong or weak?

    -   Is the relationship linear or nonlinear?

## Lab: Advice {.smaller}

-   ***Be Specific*** in your narrative.
    Don't just say "the spread is small and the center is low".
    What does that mean???
    Give numbers, units (if available), etc.

-   ***Be Specific*** in your plot labels.

-   Answer all parts of the question!!
    Statements like "Compare..." mean to do so in the written answer!

-   Written answers should be text outside of code chunks, not comments

-   Make sure you can render to PDF early!!!

## Outline {.smaller}

-   ***Last Time:*** Started learning about data transformation!

-   ***Today:***

    -   Review from last time + finish AE04

    -   More about the pipe

    -   Transformation + plotting

```{r}
#| label: slice
#| code-line-numbers: "|1|2"
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
bechdel <- read_csv("data/bechdel.csv")
```

## Quick Review: Row operations {.smaller}

::: columns
::: {.column width="70%"}
-   `slice()`: chooses rows based on location\
-   `filter()`: chooses rows based on column values\
-   `arrange()`: changes the order of the rows\
-   `sample_n()`: take a random subset of the rows\
:::

::: {.column width="30%"}
| X1  | X2  | X3  |
|:---:|:---:|:---:|
|  1  |  a  | yes |
|  3  |  b  | no  |
|  5  |  a  | yes |
|  7  |  b  | yes |
|  9  |  a  | yes |
:::
:::

## Quick Review: Column operations {.smaller}

::: columns
::: {.column width="70%"}
-   `select()`: changes whether or not a column is included\
-   `rename()`: changes the name of columns\
-   `mutate()`: changes the values of columns and creates new columns\
:::

::: {.column width="30%"}
| X1  | X2  | X3  |
|:---:|:---:|:---:|
|  1  |  a  | yes |
|  3  |  b  | no  |
|  5  |  a  | yes |
|  7  |  b  | yes |
|  9  |  a  | yes |
:::
:::

## Quick Review: Groups of rows {.smaller}

::: columns
::: {.column width="70%"}
-   `summarize()`: collapses a group into a single row\
-   `count()`: count unique values of one or more variables\
-   `group_by()`: perform calculations separately for each value of a variable\
:::

::: {.column width="30%"}
| X1  | X2  | X3  |
|:---:|:---:|:---:|
|  1  |  a  | yes |
|  3  |  b  | no  |
|  5  |  a  | yes |
|  7  |  b  | yes |
|  9  |  a  | yes |
:::
:::

# AE-04

## Recap: Group by, summarize, mutate {.smaller}

What does group by do here?

```{r}
#| label: pipe-plot-25
bechdel |>
  group_by(binary) |>
  summarise(mean_roi = mean(roi, na.rm = TRUE), 
            mean_budget = mean(budget_2013)) 
```

## Recap: Group by, summarize, mutate {.smaller}

What does group by do here?

```{r}
#| label: pipe-plot-3
bechdel |>
  #group_by(binary) |>
  summarise(mean_roi = mean(roi, na.rm = TRUE), 
            mean_budget = mean(budget_2013)) 
```

## Recap: Group by, summarize, mutate {.smaller}

What if I change summarize to mutate?

```{r}
#| label: pipe-plot-4
bechdel |>
  group_by(binary) |>
  mutate(mean_roi = mean(roi, na.rm = TRUE), 
            mean_budget = mean(budget_2013)) 
```

## Recap: Group by, summarize, mutate {.smaller}

What if I change summarize to mutate?

```{r}
#| label: pipe-plot-5
bechdel |>
  group_by(binary) |>
  mutate(mean_roi = mean(roi, na.rm = TRUE), 
            mean_budget = mean(budget_2013)) |>
  select(title, binary, mean_roi, mean_budget)
```

## Recap: Group by, summarize, mutate {.smaller}

You can group by more than one variable!

```{r}
#| label: gb-multi
bechdel |>
  group_by(binary, year) |>
  summarise(mean_roi = mean(roi, na.rm = TRUE), 
            mean_budget = mean(budget_2013)) 
```

## Recap: Group by, summarize, mutate {.smaller}

You can un group with `ungroup()`

```{r}
#| label: gb-multi-ug
bechdel |>
  group_by(binary, year) |>
  summarise(mean_roi = mean(roi, na.rm = TRUE), 
            mean_budget = mean(budget_2013)) |>
  ungroup()
```

## More about the pipe {.smaller}

-   The pipe operator passes what comes before it into the function that comes after it as the first argument in that function.

```{r}
sum(1, 2)
```

<br>

## More about the pipe {.smaller}

-   The pipe operator passes what comes before it into the function that comes after it as the first argument in that function.

```{r}
sum(1, 2)
```

<br>

```{r}
1 |> 
  sum(2)
```

## Pipe + ggplot() !! {.smaller}

```{r}
#| label: pipe-plot
ggplot(bechdel, aes(x = budget_2013)) +
  geom_boxplot()
```

## Pipe + ggplot() !! {.smaller}

```{r}
#| label: pipe-plot-2s
bechdel |>
  ggplot(aes(x = budget_2013)) +
  geom_boxplot()
```

## Why is this useful?

-   We can do data transformation immediately followed by a plot!

-   Normally, even if we are just plotting, we use the pipe with ggplot().

## Plot + data transform {.smaller}

```{r}
#| label: pipe-plot-7

bechdel
```

## Plot + data transform {.smaller}

```{r}
#| label: pipe-plot-8

bechdel |>
  mutate(budget_in_millions = budget_2013/1000000)
```

## Plot + data transform {.smaller}

```{r}
#| label: pipe-plot-9

bechdel |>
  mutate(budget_in_millions = budget_2013/1000000) |>
  ggplot(aes(x = budget_in_millions))
```

## Exploratory Data Analysis

What is exploratory data analysis (EDA)??

-   Basically everything we have done so far

-   Making plots and computing summary statistics (proportions, means, IQR, etc.) to help *explore* the data

# AE 05 {background-image="images/spirit.png" background-size="cover" background-position="center" background-repeat="no-repeat" background-opacity="0.7" style="margin-top:-16vh;"}

::: {style="position:absolute; bottom:15.9em; right:0.01em; font-size:0.4em; color:rgba(255,255,255,0.8);"}
<a href="https://your.source.link" style="color:inherit; text-decoration:underline;"> Image source </a>
:::

## Assignment {.smaller}

Let's make a tiny data frame to use as an example:

```{r}
#| message: false

library(tidyverse)
df <- tibble(x = c(1, 2, 3, 4, 5), y = c("a", "a", "b", "c", "c"))
df
```

## Assignment {.smaller}

::: columns
::: {.column .fragment width="50%"}
> **Do something and show me**

```{r}
df |>
  mutate(x = x * 2)
df
```
:::

::: {.column .fragment width="50%"}
> **Do something and save result**

```{r}
df <- df |>
  mutate(x = x * 2)
df
```
:::
:::

## Assignment {.smaller}

::: columns
::: {.column .fragment width="50%"}
> **Do something, save result, overwriting original**

```{r}
#| code-line-numbers: "5-7"

df <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c("a", "a", "b", "c", "c")
)
df <- df |>
  mutate(x = x * 2)
df
```
:::

::: {.column .fragment width="50%"}
> **Do something, save result, *not* overwriting original**

```{r}
#| code-line-numbers: "5-7"

df <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c("a", "a", "b", "c", "c")
)
df_new <- df |>
  mutate(x = x * 2)
df_new
```
:::
:::

## Assignment {.smaller}

::: columns
::: {.column .fragment width="50%"}
> **Do something, save result, overwriting original *when you shouldn't***

```{r}
#| code-line-numbers: "5-8"

df <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c("a", "a", "b", "c", "c")
)
df <- df |>
  group_by(y) |>
  summarize(mean_x = mean(x))
df
```
:::

::: {.column .fragment width="50%"}
> **Do something, save result, *not* overwriting original *when you shouldn't***

```{r}
#| code-line-numbers: "5-8"

df <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c("a", "a", "b", "c", "c")
)
df_summary <- df |>
  group_by(y) |>
  summarize(mean_x = mean(x))
df_summary
```
:::
:::

## Assignment {.smaller}

::: columns
::: {.column .fragment width="50%"}
> **Do something, save result, overwriting original**\
> ***data frame***

```{r}
#| code-line-numbers: "5-7"

df <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c("a", "a", "b", "c", "c")
)
df <- df |>
  mutate(z = x + 2)
df
```
:::

::: {.column .fragment width="50%"}
> **Do something, save result, overwriting original**\
> ***column***

```{r}
#| code-line-numbers: "5-7"

df <- tibble(
  x = c(1, 2, 3, 4, 5), 
  y = c("a", "a", "b", "c", "c")
)
df <- df |>
  mutate(x = x + 2)
df
```
:::
:::
