---
title: "Data Types and Classes"
subtitle: "Lecture 8"
date: "2026-05-26"
format: 
  revealjs:
    output-file: 08-data-types-slides.html
  html: 
    output-file: 08-data-types.html
auto-stretch: false
---

```{r}
#| echo: false
#| message: false

library(tidyverse)
library(ggthemes)
ggplot2::theme_set(theme_gray(base_size = 16))
```

# Types and classes

## Types and classes

An object's *type* indicates how it is stored in memory

## Common data types

::::: columns
::: {.column width="50%"}
You'll commonly encounter:

-   `logical`
-   `integer`
-   `double`
-   `character`
:::

::: {.column width="50%"}
You'll less commonly encounter:

-   `list`
-   `NULL`
-   `complex`
-   `raw`
:::
:::::

## Logical and Character {.smaller}

::::: columns
::: {.column width="50%"}
**Logical:** Boolean `TRUE` / `FALSE` values

-   `<dbl>`

```{r}
typeof(TRUE)
```

<br>

```{r}
typeof(FALSE)
```
:::

::: {.column width="50%"}
**Character:** single characters, or strings of characters; wrap in quotes

-   `<chr>`

```{r}
typeof("x")
```

<br>

```{r}
typeof("Hello!")
```

<br>

```{r}
typeof("TRUE")
```

:::
:::::

## Numeric: Double and integer {.smaller}

::::: columns
::: {.column width="50%"}
**Double**: floating point numerical values (default numerical type)

-   `<dbl>`

```{r}
typeof(2.5)
```

<br>

```{r}
typeof(3)
```
:::

::: {.column width="50%"}
**Integer**: integer numerical values; indicated with an `L`

-   `<int>`

```{r}
typeof(3L)
```
:::
:::::

## Type Compatibility {.smaller}

Can you use different types of data together?
Sometimes...
but be careful!

```{r}
#| error: true

"3" + 3
```

```{r}
#| error: true

3L + 3
typeof(3L + 3)
```

```{r}
#| error: true

TRUE + 3L
typeof(TRUE + 3L)
```

## Concatenation {.smaller}

**Vectors** are constructed using the `c` function

::::: columns
::: {.column width="50%"}
-   **Double** vector:

    ```{r}
    x <- c(1, 2, 3, 5)
    typeof(x)
    ```

-   **Integer** vectors:

    ```{r}
    x <- c(1L, 2L, 3L, 5L)
    typeof(x)
    ```
:::

::: {.column width="50%"}
-   **Character** vector:

    ```{r}
    x <- c("1", "2", "3", "5")
    typeof(x)
    ```

-   **Logical** vectors:

    ```{r}
    x <- c(TRUE, FALSE, FALSE)
    typeof(x)
    ```
:::
:::::

## Converting between types {.smaller}

::: hand
without intention...
:::

```{r}
c(2, "Just this one!")
```

<br>

R will happily convert between various types without complaint when different types of data are concatenated in a vector.
This is NOT always a good thing.

## Converting between types {.smaller}

::: hand
without intention...
:::

```{r}
c(FALSE, 3L)
```

<br>

```{r}
c(1.2, 3L)
```

<br>

```{r}
c(2L, "two")
```

## Converting between types

::: hand
with intention...
:::

::::: columns
::: {.column width="49%"}
```{r}
x <- 1:3
x
typeof(x)
```
:::

::: {.column .fragment width="49%"}
```{r}
y <- as.character(x)
y
typeof(y)
```
:::
:::::

## Converting between types

::: hand
with intention...
:::

::::: columns
::: {.column width="49%"}
```{r}
x <- c(TRUE, FALSE)
x
typeof(x)
```
:::

::: {.column .fragment width="49%"}
```{r}
y <- as.numeric(x)
y
typeof(y)
```
:::
:::::

## Explicit vs. implicit coercion {.smaller}

::::: columns
::: {.column width="49%"}
**Explicit coercion:**

When you call a function like:

-   `as.logical()`

-   `as.numeric()`

-   `as.integer()`

-   `as.double()`

-   `as.character()`.
:::

::: {.column .fragment width="49%"}
**Implicit coercion:**

Happens when you use a vector in a specific context that expects a certain type of vector.
:::
:::::

## You've seen explicit coercion before

```{r}
#| eval: false
#| code-line-numbers: "6"
statsci |>
  pivot_longer(
    cols = -degree_type,
    values_to = "n",
    names_to = "year",
    names_transform = as.numeric
  )
```

# Data classes

## Data classes {.smaller}

::: incremental
-   Data types are like Lego building blocks
-   We can stick them together to build more complicated constructs, e.g. *representations of data*
-   The **class** determines this construct
-   Examples: factors, dates, and data frames
:::

## Data frames {.smaller}

We can think of data frames like like vectors of equal length glued together

::::: columns
::: column
```{r}
df <- data.frame(x = 1:2, y = 3:4)
df
```

<br>

```{r}
typeof(df)
```

<br>

```{r}
class(df)
```
:::

::: column
```{r}
## alternatively, "bind" two columns together
x <- data.frame(x = 1:2)
y <- data.frame(y = 3:4)

df <- cbind(x, y)
df
```

<br>

```{r}
typeof(df)
```

<br>

```{r}
class(df)
```
:::
:::::

## Data frames {.smaller}

We can think of data frames like like vectors of equal length glued together

```{r}
df <- data.frame(x = 1:2, y = 3:4)
df
```

::::: columns
::: column
```{r}
typeof(df)
```
:::

::: column
```{r}
class(df)
```
:::
:::::

-   When we use the `pull()` function, we extract a vector from the data frame; this is functionally the same as accessing a column with  `df$col_name`

```{r}
df |>
  pull(y)
```

## Dates

```{r}
today <- as.Date("2026-05-26")
today
```

```{r}
typeof(today)
```

```{r}
class(today)
```

## More on dates

We can think of dates like an integer (the number of days since the origin, 1 Jan 1970) and an integer (the origin, aka "the Unix epoch") glued together

```{r}
as.integer(today)
```

```{r}
as.integer(today) / 365 # roughly 56 yrs
```

```{r}
as.integer(as.Date("1970-01-01"))
```

## More on dates

The `lubridate` package allows you to work with / access elements of dates seamlessly

```{r}
library(lubridate)
```
```{r}
year(today)
```

```{r}
month(today)
```

```{r}
day(today)
```


## Factors {.scrollable .smaller}

-   R uses **factors** to handle **categorical variables** with a fixed and known set of possible values

-   `factor(x = ...)`: *"The default (ordering of levels) is the unique set of values taken by as.character(x), sorted into increasing (alphabetical) order of x"*

```{r}
summer_months <- c("June", "July", "June", "August", "June")
typeof(summer_months)
summer_factor <- factor(summer_months)
summer_factor
levels(summer_factor) ## print levels
```

::::: columns
::: {.column .fragment}
```{r}
typeof(summer_factor)
```
:::

::: {.column .fragment}
```{r}
class(summer_factor)
```
:::
:::::

## More on factors

-   We can think of factors like character (level labels) and an integer (level numbers) glued together

```{r}
glimpse(summer_factor)
```

```{r}
as.integer(summer_factor)
```

## More on factors {.smaller}

We can use the **forcats** package (in **tidyverse**) to work with factors!

Some commonly used functions are:

-   fct_relevel(): reorder factors by hand

-   fct_reorder(): reorder factors by another variable

-   fct_infreq(): reorder factors by frequency

-   fct_rev(): reorder factors by reversing

## Example factor re-order

```{r}
amounts <- c("low", "medium", "high", "high", "medium")
amounts_factor <- factor(amounts)
amounts_factor
```

```{r}
fct_relevel(amounts_factor, c("low", "medium", "high"))
```

# FAQ

## Quotes VS no quotes VS backticks

. . .

```{r}
df <- tibble(
  x = c(-2, -0.5, 0.5, 1, 2),
  `2011` = c(-2, -0.5, 0.5, 1, 2),
  `my var` = c(-2, -1, 0, 1, 2)
)
df
```

## Quotes VS no quotes VS backticks {.small .scrollable}

```{r}
df <- tibble(
  x = c(-2, -0.5, 0.5, 1, 2),
  `2011` = c(-2, -0.5, 0.5, 1, 2),
  `my var` = c(-2, -1, 0, 1, 2)
)
```

Referencing a column in a pipeline:

::: columns
::: {.column .fragment width="33%"}
```{r}
df |>
  filter("x" > 0)
```
`"x"` means the literal character string.
:::

::: {.column .fragment width="31%"}
```{r}
df |>
  filter(x > 0)
```
`x` means the column name in `df`.
:::


::: {.column .fragment width="33%"}
```{r}
df |>
  filter(`x` > 0)
```
``` `x` ``` also means the column name in `df`.
:::

:::

## Quotes VS no quotes VS backticks {.small .scrollable}

```{r}
df <- tibble(
  x = c(-2, -0.5, 0.5, 1, 2),
  `2011` = c(-2, -0.5, 0.5, 1, 2),
  `my var` = c(-2, -1, 0, 1, 2)
)
```

Referencing a column in a pipeline:

::: columns
::: {.column .fragment width="33%"}
```{r}
df |>
  filter("2011" > 0)
```
`"2011"` means the literal character string.
:::

::: {.column .fragment width="31%"}
```{r}
df |>
  filter(2011 > 0)
```
`2011` means the literal number.
:::

::: {.column .fragment width="33%"}
```{r}
df |>
  filter(`2011` > 0)
```
``` `2011` ``` means the column name in `df`.
:::

:::

## Quotes VS no quotes VS backticks {.small .scrollable}

```{r}
df <- tibble(
  x = c(-2, -0.5, 0.5, 1, 2),
  `2011` = c(-2, -0.5, 0.5, 1, 2),
  `my var` = c(-2, -1, 0, 1, 2)
)
```

Referencing a column in a pipeline:

::: columns
::: {.column .fragment width="33%"}
```{r}
df |>
  filter("my var" > 0)
```
`"my var"` means the literal character string.
:::

::: {.column .fragment width="31%"}
```{r}
#| error: true
df |>
  filter(my var > 0)
```
`my var` means nothing.
:::

::: {.column .fragment width="33%"}
```{r}
df |>
  filter(`my var` > 0)
```
``` `my var` ``` means the column name in `df`.
:::

:::

# AE 08: Working with Factors
