---
title: "Multiple Linear Regression 1"
subtitle: "Lecture 14"
date: "2026-06-05"
format: 
  revealjs:
    output-file: 14-mlr-I-slides.html
  html: 
    output-file: 14-mlr-I.html
auto-stretch: false
---

```{r}
#| include: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(scatterplot3d)
library(ggthemes)
library(ggbeeswarm)
```

# Recap: simple linear regression

## From last time...penguins

::: question
A researcher wants to see how body mass varies with flipper length.
:::

. . .

-   outcome: body mass (g) (numerical)

-   predictor: flipper length (mm) (numerical)

. . .

Flipper length is easier to measure, so more plausible you would predict body mass based on that and not the other way around.

## How do we concisely summarize the association between two variables?

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4
#| out-width: "65%"
#| fig-align: center

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point()
```

## Line of best fit!

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4
#| out-width: "65%"
#| fig-align: center

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "cornflowerblue")
```

## Concise numerical summary: correlation

```{r}
penguins |>
  summarize(
    r = cor(flipper_length_mm, body_mass_g, use = "complete.obs")
  )
```

Measures the strength and direction of the linear association between two numerical variables.
Strong and positive in this case.

## Model fit

```{r}
bm_fl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm, data = penguins)

tidy(bm_fl_fit)
```

$$
\widehat{\text{body mass (g)}}=-5780.83+49.68\times \text{flipper length (mm)}
$$

## $R^2$ (coefficient of determination)

The fraction of the variation in the response explained by the model.
A number between 0 (bad) and 1 (good) that measures goodness-of-fit:

::::: columns
::: {.column width="50%"}
```{r}
bm_fl_fit |>
  glance() |>
  select(r.squared)
```
:::

::: {.column width="50%"}
```{r}
penguins |>
  summarize(
    r = cor(flipper_length_mm, body_mass_g, use = "complete.obs"),
    R2 = r^2
  )
```
:::
:::::

## Interpretation

$$
\widehat{\text{body mass (g)}}=-5780.83+49.68\times \text{flipper length (mm)}
$$

-   Intercept: Penguins with 0 mm flipper length are expected, on average, to weigh -5,781 grams (makes no sense!);
-   Slope: For each additional mm of a penguin's flipper length, the weight of the penguin is expected to be higher, on average, by 49.7 g.

## Prediction

```{r}
#| echo: false
#| warning: false

new_penguin <- tibble(flipper_length_mm = 205)

preds <- predict(bm_fl_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "cornflowerblue")   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "deeppink"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "deeppink"
  )
```

## Prediction

```{r}
#| label: predict-210
new_penguin <- tibble(flipper_length_mm = 205)

predict(bm_fl_fit, new_data = new_penguin)
```

$$
\widehat{\text{body mass (g)}}=-5780.83+49.68\times 205\approx 4404.71\text{g}
$$

. . .

We predict that a penguin whose flipper is 205 mm long will weigh 4,404.71 g, on average.

# Linear regression with one categorical predictor

## New question

::: question
A different researcher wants to look at body weight of penguins based on the island they were recorded on.
How are the variables involved in this analysis different?
:::

. . .

-   outcome: body mass in grams (numerical)

-   predictor: island (categorical, with three levels)

. . .

```{r}
levels(penguins$island)
```

## Scatterplot: possible, but not so good {.smaller}

::: task
Visualize the relationship between body weight and island of penguins.
Also calculate the average body weight per island.
:::

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4
#| out-width: "55%"
#| fig-align: center

ggplot(penguins, aes(x = island, y = body_mass_g)) + 
  geom_point()
```

## Boxplot {.smaller}

::: question
Visualize the relationship between body weight and island of penguins.
Also, calculate the average body weight per island.
:::

```{r}
#| output-location: column
#| warning: false
ggplot(
  penguins,
  aes(x = island, y = body_mass_g)
) +
  geom_boxplot() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body mass by island"
  )
```

```{r}
#| output-location: column
penguins |>
  group_by(island) |>
  summarize(
    bm_mean = mean(body_mass_g, na.rm = TRUE)
  )
```

## Density plot {.smaller}

```{r}
#| output-location: column
#| warning: false
ggplot(
  penguins,
  aes(x = body_mass_g, color = island, fill = island)
) +
  geom_density(alpha = 0.5) +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body Mass by Island",
    color = "Island",
    fill = "Island"
  )
```

## Violin plot {.smaller}

```{r}
#| output-location: column
#| warning: false
ggplot(
  penguins,
  aes(x = island, y = body_mass_g)
) +
  geom_violin() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body Mass by Island"
  )
```

## Multiple geoms {.smaller}

```{r}
#| output-location: column
#| warning: false
#| fig-asp: 0.6
ggplot(
  penguins,
  aes(x = island, y = body_mass_g, color = island)
) +
  geom_boxplot() +
  geom_beeswarm(size = 2, alpha = 0.5) +
  scale_color_colorblind() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body Mass by Island",
    color = "Island"
  ) +
  theme(legend.position = "none")
```

## Model - fit {.smaller}

::: task
Fit a linear regression model predicting body weight from island and display the results.
Why is Biscoe not on the output?
:::

```{r}
bm_island_fit <- linear_reg() |>
  fit(body_mass_g ~ island, data = penguins)

tidy(bm_island_fit)
```

. . .

Huh?

## Dummy variables {.smaller}

```{r}
dummy_penguins <- penguins |>
  select(body_mass_g, island) |>
  arrange(body_mass_g) |>
  mutate(
    islandDream = if_else(island == "Dream", 1, 0),
    islandTorgersen = if_else(island == "Torgersen", 1, 0),
  )

dummy_penguins
```

## Dummy variables - fit {.scrollable}

```{r}
linear_reg() |>
  fit(
    body_mass_g ~ islandDream + islandTorgersen, 
    dummy_penguins
  ) |>
  tidy()
```

. . .

Exactly the same as before!

## Model - interpret {.small}

$$
\widehat{\text{body mass(g)}} = 4716 - 1003 \times \text{islandDream} - 1010 \times \text{islandTorgersen}
$$

-   Intercept: [Penguins from Biscoe island are expected to weigh, on average, 4,716 g.]{.fragment}

-   Slope - islandDream: [Penguins from Dream island are expected to weigh, on average, 1,003 g **less than those from Biscoe island**.]{.fragment}

-   Slope - islandTorgersen: [Penguins from Torgersen island are expected to weigh, on average, 1,010 g **less than those from Biscoe island**.]{.fragment}

## Model - predict {.smaller}

::: task
What is the predicted body weight of a penguin on Biscoe island?
What are the estimated body weights of penguins on Dream and Torgersen islands?
Where have we seen these values before?
:::

```{r}
new_penguins = tibble(
  island = c("Biscoe", "Dream", "Torgersen")
)

predict(bm_island_fit, new_data = new_penguins)
```

## Model - predict {.small .scrollable}

::: task
Calculate the predicted body weights of penguins on Biscoe, Dream, and Torgersen islands *by hand*.
:::

$$
\widehat{body~mass} = 4716 - 1003 \times islandDream - 1010 \times islandTorgersen
$$

. . .

-   Biscoe: $\widehat{body~mass} = 4716 - 1003 \times 0 - 1010 \times 0 = 4716$

. . .

-   Dream: $\widehat{body~mass} = 4716 - 1003 \times 1 - 1010 \times 0 = 3713$

. . .

-   Torgersen: $\widehat{body~mass} = 4716 - 1003 \times 0 - 1010 \times 1 = 3706$

## Look familiar?

```{r}
penguins |>
  group_by(island) |>
  summarize(
    bm_mean = mean(body_mass_g, na.rm = TRUE)
  )
```

## Models with categorical predictors

::: incremental
-   When the categorical predictor has many levels, they're encoded as **dummy variables**.

-   The first level of the categorical variable is the "baseline" level.
    In a model with one categorical predictor, the intercept is the predicted value of the outcome for the baseline level (x = 0).

-   Each slope coefficient describes the difference between the predicted value of the outcome for that level of the categorical variable **compared to the baseline level**.
:::

## In general {.small .scrollable}

Predicting continuous outcome $Y$ using one categorical predictor $X$ with multiple levels 1, 2, ..., $k$.
Create dummy variables for every level *except* the base level:

. . .

$$
cat_k=\begin{cases}
1 & X=k\\
0 & \text{else}
\end{cases}
$$

. . .

Then fit a regression with *multiple* dummy predictors:

$$
\widehat{Y} = b_0 + b_1 \times cat_1 + b_2 \times cat_2 \ldots + b_{k-1} \times cat_{k-1}
$$

::: incremental
-   $b_0$ : the model prediction for a member of the base level;

-   $b_1$: how does the prediction change when we move from the base level to level 1?

-   $b_2$: how does the prediction change when we move from the base level to level 2?

-   etc...
:::

. . .

::: callout-important
## We're not animals. We have technology!

The computer handles *all* of this for you, but you need to understand the details so you code and interpret it correctly.
:::

## Changing the baseline level

-   By default, R uses the first level of a categorical variable as the baseline level.
    this is often the first alphabetically, but make sure you check!

-   We can change the baseline level by **reordering the levels** of the categorical variable... do you remember how to do this??

# Linear regression with one numerical and one categorical predictor

## Three variables on one 2D plot

```{r}
#| include: false 
penguins <- palmerpenguins::penguins
```

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4
#| out-width: "65%"
#| fig-align: center

ggplot(penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point()
```

## How do we predict using more than one predictor? {.small}

Both of these models use `flipper_length_mm` *and* `island` to predict `body_mass_g`:

```{r}
#| label: additive-interaction-viz
#| layout-ncol: 2
#| echo: false
#| fig-asp: 1
#| warning: false

# Plot A
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)
bm_fl_island_aug <- augment(bm_fl_island_fit, new_data = penguins)
ggplot(
  bm_fl_island_aug, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(title = "Plot A - Additive model") +
  theme(legend.position = "bottom")

# Plot B
ggplot(
  penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Plot B - Interaction model") +
  theme(legend.position = "bottom")
```

## The additive model: parallel lines, one for each island {.small .scrollable}

```{r}
#| label: additive-viz
#| echo: false
#| warning: false

# Plot A
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)
bm_fl_island_aug <- augment(bm_fl_island_fit, new_data = penguins)
ggplot(
  bm_fl_island_aug, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(title = "Additive model") +
  theme(legend.position = "bottom")


```

## The additive model: parallel lines, one for each island {.small .scrollable}

```{r}
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)

tidy(bm_fl_island_fit)
```

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -4625 &+ 44.5 \times {\text{flipper length (mm)}} \\
&- 262 \times \text{Dream} \\
&- 185 \times \text{Torgersen}
\end{aligned}
$$

## Where do the three lines come from? {.smaller .scrollable}

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -4625 &+ 44.5 \times \text{flipper length (mm)} \\
&- 262 \times \text{Dream} \\
&- 185 \times \text{Torgersen}
\end{aligned}
$$

. . .

If penguin is from Biscoe, Dream = 0 **and** Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -4625 &+ 44.5 \times \text{flipper length (mm)}
\end{aligned}
$$

. . .

If penguin is from Dream, Dream = 1 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -4887 &+ 44.5 \times \text{flipper length (mm)}
\end{aligned}
$$

. . .

If penguin is from Torgersen, Dream = 0 and Torgersen = 1:

. . .

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -4810 &+ 44.5 \times \text{flipper length (mm)}
\end{aligned}
$$

. . .

Either way, same slope, so the lines are parallel.

## The interaction model: different lines for each island {.medium .scrollable}

```{r}
#| label: interaction-viz
#| code-fold: true
#| warning: false

ggplot(
  penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Interaction model") +
  theme(legend.position = "bottom")
```

## The interaction model: different lines for each island {.smaller .scrollable}

```{r}
bm_fl_island_int_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm * island, data = penguins)

tidy(bm_fl_island_int_fit) |> select(term, estimate)
```

. . .

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -5464 &+ 48.5 \times \text{flipper length (mm)} \\
&+ 3551 \times \text{Dream} \\
&+ 3218 \times \text{Torgersen} \\
&- 19.4 \times \text{flipper length (mm)}*\text{Dream} \\
&- 17.4 \times \text{flipper length (mm)}*\text{Torgersen}
\end{aligned}
$$

## Where do the three lines come from? {.smaller .scrollable}

$$
\begin{aligned}
\small\widehat{\text{body mass (g)}} = -5464 &+ 48.5 \times \text{flipper length (mm)} \\
&+ 3551 \times \text{Dream} \\
&+ 3218 \times \text{Torgersen} \\
&- 19.4 \times \text{flipper length (mm)} * \text{Dream} \\
&- 17.4 \times \text{flipper length (mm)} * \text{Torgersen}
\end{aligned}
$$

. . .

If penguin is from Biscoe, Dream = 0 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{\text{body mass (g)}} = -5464 &+ 48.5 \times \text{flipper length (mm)}
\end{aligned}
$$

. . .

If penguin is from Dream, Dream = 1 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{\text{body mass (g)}} &= (-5464 + 3551) + (48.5-19.4) \times \text{flipper length (mm)}\\
&=-1913+29.1\times \text{flipper length (mm)}.
\end{aligned}
$$

## Prediction on Biscoe island

```{r}
#| echo: false
#| warning: false

new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Biscoe"
  )

preds <- predict(bm_fl_island_int_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = island)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "#FE5D26"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "#FE5D26"
  )
```

## Prediction on Biscoe island

```{r}
new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Biscoe"
)

predict(bm_fl_island_int_fit, new_data = new_penguin)
```

$$
\widehat{\text{body mass (g)}} = -5464 + 48.5 \times 205
$$

## Prediction on Dream island

```{r}
#| echo: false
#| warning: false

new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Dream"
  )

preds <- predict(bm_fl_island_int_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = island)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "forestgreen"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "forestgreen"
  )
```

## Prediction on Dream island

```{r}
new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Dream"
)

predict(bm_fl_island_int_fit, new_data = new_penguin)
```

$$
\widehat{\text{body mass (g)}} = (-5464 + 3551) + (48.5 - 19.4) \times 205
$$

## Prediction on Torgersen island

```{r}
#| echo: false
#| warning: false

new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Torgersen"
  )

preds <- predict(bm_fl_island_int_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = island)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "blue"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "blue"
  )
```

## Prediction on Torgersen island 

```{r}
new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Torgersen"
)

predict(bm_fl_island_int_fit, new_data = new_penguin)
```

$$
\widehat{\text{body mass (g)}} = (-5464 + 3218) + (48.5 - 17.4) \times 205
$$

# Linear regression with many numerical predictors

## Multiple numerical predictors {.smaller .scrollable}

```{r}
bm_fl_bl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, data = penguins)

tidy(bm_fl_bl_fit)
```

. . .

$$
\small\widehat{\text{body mass (g)}}=-5736+48.1\times \text{flipper length (mm)}+6\times \text{bill length (mm)}
$$

## Interpretation {.scrollable .smaller}

$$
\small\widehat{\text{body mass (g)}}=-5736+48.1\times \text{flipper length (mm)}+6\times \text{bill length (mm)}
$$

. . .

**Interpretations**:

::: incremental
-   We predict that the body mass of a penguin with *zero* flipper length and *zero* bill length will be -5,736 g, on average (**makes no sense**);
-   Holding all other variables constant, for every additional mm in flipper length, we expect the body mass of penguins to be higher, on average, by 48.1 g;
-   Holding all other variables constant, for every additional mm in bill length, we expect the body mass of penguins to be higher, on average, by 6 g.
:::

## Prediction

```{r}
new_penguin <- tibble(
  flipper_length_mm = 200,
  bill_length_mm = 45
)

predict(bm_fl_bl_fit, new_data = new_penguin)
```

$$
\widehat{\text{body mass (g)}}=-5736+48.1\times 200+6\times 45
$$

## Picture? It's not pretty...

2 predictors + 1 response = 3 dimensions.
Ick!

```{r}
#| echo: false
#| warning: false

scatterplot3d(
  penguins[,c("bill_length_mm", "flipper_length_mm", "body_mass_g")],
  pch = 19, 
  color = "cornflowerblue",
  xlab = "Bill length (mm)", 
  ylab = "Flipper length (mm)", 
  zlab = "Body mass (grams)"
) 
```

## Picture? It's not pretty...

Instead of a line of best fit, it's a *plane* of best fit.
Double ick!

```{r}
#| echo: false
#| warning: false

plot3vars <- scatterplot3d(
  penguins[,c("bill_length_mm", "flipper_length_mm", "body_mass_g")],
  pch = 19, 
  color = "cornflowerblue",
  xlab = "Bill length (mm)", 
  ylab = "Flipper length (mm)", 
  zlab = "Body mass (grams)"
) 

model <- lm(body_mass_g ~ bill_length_mm + flipper_length_mm, data = penguins)
plot3vars$plane3d(model, col = "purple")
```

# Multiple linear regression in general

## Population model {.smaller}

**Multiple linear regression** captures the relationship between a numerical outcome $Y$ and *many* numerical predictors $X_1$, $X_2$, ..., $X_p$:

$$\Large{Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + ...+ \beta_p X_p+\epsilon}$$

::: incremental
-   $\beta_0$: True intercept of the relationship between the $X_j$ and $Y$
-   $\beta_j$: True slope of the relationship between $X_j$ and $Y$
-   $\epsilon$: Error (residual)
:::

. . .

The model with the greek letters and the error term is the "true," idealized, *population* relationship that we could access if we had infinite amounts of perfect data.
But we don't, so we have to settle for...

## Estimated / fitted model {.smaller}

$$\Large{\widehat{Y} = b_0 + b_1 X_1 + b_2 X_2 + ... + b_pX_p}$$

-   $b_j$: Estimated slope of the relationship between $X_j$ and $Y$;
-   $b_0$: your prediction for $Y$ if all the predictors equal zero;
-   No error term!

. . .

This is your *best guess* at the true regression function based on the noisy, meager, imperfect data you actually have access to.
We still compute the $b_j$ using the principle of least squares: pick the estimates that make the sum of squared residuals as small as possible.

## Next time

Today we saw multiple models that are all attempting to do the same thing: predict body mass.

-   Which one is "best"?
-   What does "best" even mean?
