---
title: "Statistical inference"
subtitle: "Lecture 22"
date: "2025-6-17"
format: 
  revealjs:
    output-file: 22-more-inference-slides.html
    pdf-separate-fragments: true
auto-stretch: false
---

## While you wait... {.smaller}

```{r}
#| label: load-packages
#| message: false
#| echo: false
library(tidyverse)
library(tidymodels)
library(openintro)
library(glue)
todays_ae <- "ae-18-inference-practice"
```

::: appex
-   Go to your `ae` project in RStudio.

-   Make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   Click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Wait until you're prompted to work on the application exercise during class before editing the file.
:::

# Overview: statistical inference

## The population model (greek letters, no hats) {.smaller}

::: incremental
-   This is an *idealized* representation of the data: $$
    y = \beta_0+\beta_1x +\varepsilon;
    $$

-   There is some "true line" floating out there with a "true slope" $\beta_1$ and "true intercept" $\beta_0$;

-   With infinite amounts of perfectly measured data, we could know $\beta_0$ and $\beta_1$ exactly;

-   We don't have that, so we must use finite amounts of imperfect data to *estimate*;

-   We are especially interested in $\beta_1$ because is characterizes the association between $x$ and $y$, which is useful for *prediction*.

-   This idea continues to multiple linear regression (additive/interaction) and logistic regression
:::

## "Three branches of statistical government" {.smaller}

We have an unknown quantity we are trying to learn about (for example, $\beta_1$) using noisy, imperfect data.
Learning comes in three flavors:

::: incremental
-   **POINT ESTIMATION**: get a single-number best guess for $\beta_1$;

-   **INTERVAL ESTIMATION**: get a range of likely values for $\beta_1$ that characterizes (sampling) uncertainty;

-   **HYPOTHESIS TESTING**: use the data to distinguish competing claims about $\beta_1$.
:::

. . .

::: callout-important
While the examples we do here follow simple linear regression, the same ideas apply to the other models we have learned so far!!
:::

## Point estimation (roman letters, hats!) {.smaller}

We estimate $\beta_0$ and $\beta_1$ with the coefficients of the best fit line:

$$
\hat{y}=b_0+b_1x.
$$

-   Linear regression: "Best" means "least squares." We pick the estimates so that the sum of squared residuals is as small as possible.

## Point estimation {.smaller}

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

## Point estimation {.smaller}

```{r}
linear_reg() |>
  fit(mpg ~ wt, data = mtcars) |>
  tidy()
```

## Sampling uncertainty {.smaller}

::: incremental
-   How do our point estimates vary across alternative, hypothetical datasets?

    -   If they vary alot, uncertainty is high, and results are unreliable;
    -   if they vary only a little, uncertainty is low, and results are more reliable.

-   We can use the *bootstrap* to construct alternative datasets and assess the sensitivity of our estimates to changes in the data.
:::

## Bootstrap distribution {.smaller}

This histogram displays variation in the slope estimate across alternative datasets.

```         
- Large spread >> high uncertainty;
- Small spread >> lower uncertainty.
```

```{r}
#| echo: false

df_fit <- linear_reg() |>
  fit(price ~ area, data = duke_forest)

intercept <- tidy(df_fit) |> filter(term == "(Intercept)") |> pull(estimate) |> round()
slope <- tidy(df_fit) |> filter(term == "area") |> pull(estimate) |> round()

set.seed(119)

df_boot_samples_100 <- duke_forest |>
  specify(price ~ area) |>
  generate(reps = 100, type = "bootstrap")

df_boot_samples_100_fit <- df_boot_samples_100 |>
  fit()

df_boot_samples_100_hist <- ggplot(df_boot_samples_100_fit |> filter(term == "area"), aes(x = estimate)) +
  geom_histogram(binwidth = 10, color = "white") +
  geom_vline(xintercept = slope, color = "#8F2D56", linewidth = 1) +
  labs(x = "Slope", y = "Count",
       title = "Slopes of 100 bootstrap samples") +
  scale_x_continuous(labels = label_dollar())

df_boot_samples_100_hist
```

## Confidence interval {.smaller}

Pick a range that swallows up a large % of the histogram:

```{r}
#| echo: false
lower <- df_boot_samples_100_fit |>
  ungroup() |>
  filter(term == "area") |>
  summarise(quantile(estimate, 0.025)) |>
  pull()

upper <- df_boot_samples_100_fit |>
  ungroup() |>
  filter(term == "area") |>
  summarise(quantile(estimate, 0.975)) |>
  pull()

df_boot_samples_100_hist +
  geom_vline(xintercept = lower, color = "steelblue", linewidth = 1, linetype = "dashed") +
  geom_vline(xintercept = upper, color = "steelblue", linewidth = 1, linetype = "dashed")
```

We use quantiles (think IQR) but there are other ways.

## Hypothesis testing {.smaller}

::: incremental
-   Two competing claims about $\beta_1$: $$
    \begin{aligned}
    H_0&: \beta_1=0\quad(\text{nothing going on})\\
    H_A&: \beta_1\neq0\quad(\text{something going on})
    \end{aligned}
    $$

-   Do the data strongly favor one or the other?

-   How can we quantify this?
:::

## Hypothesis testing {.smaller}

::: incremental
-   Think hypothetically: if the null hypothesis were in fact true, would my results be out of the ordinary?

    -   if no, then the null could be true;
    -   if yes, then the null might be bogus;

-   My results represent the reality of actual data.
    If they conflict with the null, then you throw out the null and stick with reality;

-   How do we *quantify* "would my results be out of the ordinary"?
:::

## Null distribution {.smaller}

If the null happened to be true, how would we expect our results to vary across datasets?
We can use simulation to answer this:

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4

set.seed(34)
fake_draws <- tibble(
  replicate = 1:500,
  x = rnorm(500),
  stat = x,
  estimate = x,
) 

fake_draws |>
  ggplot(aes(x = x)) + 
  geom_histogram() + 
  labs(
    title = "Null distribution: variation in the slope estimate if the null were true",
    x = "hypothetical slope estimates",
    y = "count"
  ) + 
  xlim(-5, 5)
```

This is how the world should look *if the null is true*.

## Null distribution versus reality {.smaller}

Locate the *actual* results of your *actual* data analysis under the null distribution.
Are they in the middle?
Are they in the tails?

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4

fake_draws |>
  ggplot(aes(x = x)) + 
  geom_histogram() + 
  geom_vline(xintercept = 0.5, color = "red", size = 2) + 
  annotate("text", x = 2, y = 65, label = "Your actual\nslope estimate", color = "red", size = 3*.pt) + 
  labs(
    title = "Null distribution: variation in the slope estimate if the null were true",
    x = "hypothetical slope estimates",
    y = "count"
  ) + 
  xlim(-5, 5)
```

Are these results in harmony or conflict with the null?

## Null distribution versus reality {.smaller}

Locate the *actual* results of your *actual* data analysis under the null distribution.
Are they in the middle?
Are they in the tails?

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4

fake_draws |>
  ggplot(aes(x = x)) + 
  geom_histogram() + 
  geom_vline(xintercept = 5, color = "red", size = 2) + 
  annotate("text", x = 3.5, y = 65, label = "Your actual\nslope estimate", color = "red", size = 3*.pt) + 
  labs(
    title = "Null distribution: variation in the slope estimate if the null were true",
    x = "hypothetical slope estimates",
    y = "count"
  ) + 
  xlim(-5, 5)
```

Are these results in harmony or conflict with the null?

## Null distribution versus reality {.smaller}

Locate the *actual* results of your *actual* data analysis under the null distribution.
Are they in the middle?
Are they in the tails?

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4

fake_draws |>
  ggplot(aes(x = x)) + 
  geom_histogram() + 
  geom_vline(xintercept = 1.75, color = "red", size = 2) + 
  annotate("text", x = 3.5, y = 65, label = "Your actual\nslope estimate", color = "red", size = 3*.pt) + 
  labs(
    title = "Null distribution: variation in the slope estimate if the null were true",
    x = "hypothetical slope estimates",
    y = "count"
  ) + 
  xlim(-5, 5)
```

Are these results in harmony or conflict with the null?

## *p*-value {.smaller}

::: incremental
The $p$-value is the probability of being *even farther out* in the tails of the null distribution than your results already were.

-   if this number is very low, then your results would be out of the ordinary if the null were true, so maybe the null was never true to begin with;

-   if this number is high, then your results may be perfectly compatible with the null.
:::

## *p*-value {.smaller}

*p*-value is the fraction of the histogram area shaded red:

```{r}
#| echo: false
#| message: false
#| warning: false
set.seed(3)
x <- rnorm(100, mean = 0.1, sd = 2)
df <- tibble(x = x)
obs_stat <- df |>
  specify(response = x) |>
  calculate(stat = "mean")
null_dist <- df |>
  specify(response = x) |>
  hypothesize(null = "point", mu = 0) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean")


visualize(null_dist) + 
  shade_p_value(obs_stat = obs_stat, direction = "two-sided")
```

. . .

Big ol' *p*-value.
Null looks plausible

## *p*-value {.smaller}

*p*-value is the fraction of the histogram area shaded red:

```{r}
#| echo: false
#| message: false
#| warning: false
set.seed(3)
x <- rnorm(100, mean = 1, sd = 2)
df <- tibble(x = x)
obs_stat <- df |>
  specify(response = x) |>
  calculate(stat = "mean")
null_dist <- df |>
  specify(response = x) |>
  hypothesize(null = "point", mu = 0) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean")


visualize(null_dist) + 
  shade_p_value(obs_stat = obs_stat, direction = "two-sided")
```

. . .

*p*-value is basically zero.
Null looks bogus.

## *p*-value {.smaller}

*p*-value is the fraction of the histogram area shaded red:

```{r}
#| echo: false
#| message: false
#| warning: false
set.seed(3)
x <- rnorm(100, mean = 0.5, sd = 4)
df <- tibble(x = x)
obs_stat <- df |>
  specify(response = x) |>
  calculate(stat = "mean")
null_dist <- df |>
  specify(response = x) |>
  hypothesize(null = "point", mu = 0) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean")


visualize(null_dist) + 
  shade_p_value(obs_stat = obs_stat, direction = "two-sided")
```

. . .

*p*-value is...kinda small?
Null looks...?

## Discernibility level {.smaller}

::: incremental
-   How do we decide if the *p*-value is big enough or small enough?

-   Pick a threshold $\alpha\in[0,\,1]$ called the *discernibility level*:

    -   If $p\text{-value} < \alpha$, reject null and accept alternative;
    -   If $p\text{-value} \geq \alpha$, fail to reject null;

-   Standard choices: $\alpha=0.01, 0.05, 0.1, 0.15$.
:::

## Pop quiz! (not really)

We will do a few multiple choice questions to check-in about statistical inference:

## What is $b_1$? {.smaller}

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4
set.seed(1)
n = 50
x <- scale(rnorm(n))[,1]
y <- 3 + 0.5 * x + rnorm(n)
my_model <- lm(y ~ x)
x <- cor(x, y) * sd(y) * x / 0.5
y <- y - (my_model$coefficients[1] - 3)
df <- tibble(x = x, y = y)

ggplot(df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(axis.text = element_text(size = 20))
```

::: columns
::: column
a.  -1 / 2
b.  0
c.  1 / 4
:::

::: column
d.  1 / 2
e.  1
f.  2
:::
:::

## What is $b_1$? {.smaller}

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4
set.seed(1)
n = 50
x <- scale(rnorm(n))[,1]
y <- 3 + 0.5 * x + rnorm(n)
my_model <- lm(y ~ x)
x <- cor(x, y) * sd(y) * x / 0.5
y <- y - (my_model$coefficients[1] - 3)
df <- tibble(x = x, y = y)

ggplot(df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(axis.text = element_text(size = 20))
```

::: columns
::: column
a.  -1 / 2
b.  0
c.  1 / 4
:::

::: column
d.  1 / 2 (*correct*)
e.  1
f.  2
:::
:::

## What is $b_0$? {.smaller}

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4
set.seed(1)
n = 50
x <- scale(rnorm(n))[,1]
y <- 3 + 0.5 * x + rnorm(n)
my_model <- lm(y ~ x)
x <- cor(x, y) * sd(y) * x / 0.5
y <- y - (my_model$coefficients[1] - 3)
df <- tibble(x = x, y = y)

ggplot(df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(axis.text = element_text(size = 20))
```

::: columns
::: column
a.  1
b.  1.75
:::

::: column
c.  2
d.  3
:::
:::

## What is $b_0$? {.smaller}

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4
set.seed(1)
n = 50
x <- scale(rnorm(n))[,1]
y <- 3 + 0.5 * x + rnorm(n)
my_model <- lm(y ~ x)
x <- cor(x, y) * sd(y) * x / 0.5
y <- y - (my_model$coefficients[1] - 3)
df <- tibble(x = x, y = y)

ggplot(df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(axis.text = element_text(size = 20))
```

::: columns
::: column
a.  1
b.  1.75
:::

::: column
c.  2
d.  3 (*correct*)
:::
:::

## Which of these could be a bootstrap sample? {.smaller}

This is the original dataset:

```{r}
#| echo: false
#| message: false
#| warning: false

set.seed(3456789)
original_dataset <- tibble(scores = round(rnorm(5), 3))
glimpse(original_dataset)
```

| A      |     |     | B      |     |     | C      |     |     | D      |
|--------|-----|-----|--------|-----|-----|--------|-----|-----|--------|
| -0.515 |     |     | 1.563  |     |     | -0.515 |     |     | -0.522 |
| -0.411 |     |     | -0.411 |     |     | -0.411 |     |     | 1.12   |
| 1.563  |     |     | -0.515 |     |     | -0.411 |     |     | 1.206  |
| -0.515 |     |     |        |     |     | 1.563  |     |     | 0.68   |
| -0.515 |     |     |        |     |     | 1.563  |     |     | 0.83   |
|        |     |     |        |     |     | 0.523  |     |     |        |
|        |     |     |        |     |     | 1.563  |     |     |        |

## Which of these could be a bootstrap sample? {.smaller}

This is the original dataset:

```{r}
#| echo: false
#| message: false
#| warning: false

set.seed(3456789)
original_dataset <- tibble(scores = round(rnorm(5), 3))
glimpse(original_dataset)
```

| A (*correct*) |     |     | B      |     |     | C      |     |     | D      |
|---------------|-----|-----|--------|-----|-----|--------|-----|-----|--------|
| -0.515        |     |     | 1.563  |     |     | -0.515 |     |     | -0.522 |
| -0.411        |     |     | -0.411 |     |     | -0.411 |     |     | 1.12   |
| 1.563         |     |     | -0.515 |     |     | -0.411 |     |     | 1.206  |
| -0.515        |     |     |        |     |     | 1.563  |     |     | 0.68   |
| -0.515        |     |     |        |     |     | 1.563  |     |     | 0.83   |
|               |     |     |        |     |     | 0.523  |     |     |        |
|               |     |     |        |     |     | 1.563  |     |     |        |

## Match the picture to the p-value. {.smaller}

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.3
#| fig-align: center

set.seed(23456)

v <- sample(c(4, 1, 0.5, 2), 4, replace = FALSE)

par(mfrow = c(1, 4), mar = c(3, 0.5, 3, 0.5))
normTail(0, 1, xlim = c(-4, 4), main = "Plot 1", U = v[1], L = -v[1])
abline(v = v[1], lty = 2, col = "red")
normTail(0, 1, xlim = c(-4, 4), main = "Plot 2", U = v[2], L = -v[2])
abline(v =v[2], lty = 2, col = "red")
normTail(0, 1, xlim = c(-4, 4), main = "Plot 3", U = v[3], L = -v[3])
abline(v = v[3], lty = 2, col = "red")
normTail(0, 1, xlim = c(-4, 4), main = "Plot 4", U = v[4], L = -v[4])
abline(v = v[4], lty = 2, col = "red")
```

::: columns
::: column
a.  0
b.  0.045
:::

::: column
c.  0.32
d.  0.62
:::
:::

## Match the picture to the p-value. {.smaller}

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.3
#| fig-align: center

set.seed(23456)

v <- sample(c(4, 1, 0.5, 2), 4, replace = FALSE)

par(mfrow = c(1, 4), mar = c(3, 0.5, 3, 0.5))
normTail(0, 1, xlim = c(-4, 4), main = "Plot 1", U = v[1], L = -v[1])
abline(v = v[1], lty = 2, col = "red")
normTail(0, 1, xlim = c(-4, 4), main = "Plot 2", U = v[2], L = -v[2])
abline(v =v[2], lty = 2, col = "red")
normTail(0, 1, xlim = c(-4, 4), main = "Plot 3", U = v[3], L = -v[3])
abline(v = v[3], lty = 2, col = "red")
normTail(0, 1, xlim = c(-4, 4), main = "Plot 4", U = v[4], L = -v[4])
abline(v = v[4], lty = 2, col = "red")
```

::: columns
::: column
a.  0 (Plot 2)
b.  0.045 (Plot 4)
:::

::: column
c.  0.32 (Plot 3)
d.  0.62 (Plot 1)
:::
:::

## Which could be the null distribution of the test? {.smaller}

For these hypotheses

$$
\begin{aligned}
H_0&: \beta_1=5\\
H_A&: \beta_1\neq 5.
\end{aligned}
$$

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4
n = 500

set.seed(1)
df <- tibble(
  draws = c(rnorm(n, mean = 5), 
            rnorm(n, mean = 0), 
            rnorm(n, mean = -5), 
            rnorm(n, mean = 2)),
  id = c(rep("A", n),
         rep("B", n),
         rep("C", n),
         rep("D", n))
)

ggplot(df, aes(x = draws)) + 
  geom_histogram() + 
  facet_wrap(~ id) + 
  labs(
    x = NULL,
    y = NULL
  )
```

## Which could be the null distribution of the test? {.smaller}

For these hypotheses

$$
\begin{aligned}
H_0&: \beta_1=5\\
H_A&: \beta_1\neq 5.
\end{aligned}
$$

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 0.4
n = 500
set.seed(1)

df <- tibble(
  draws = c(rnorm(n, mean = 5), 
            rnorm(n, mean = 0), 
            rnorm(n, mean = -5), 
            rnorm(n, mean = 2)),
  id = c(rep("A (correct)", n),
         rep("B", n),
         rep("C", n),
         rep("D", n))
)

ggplot(df, aes(x = draws)) + 
  geom_histogram() + 
  facet_wrap(~ id) + 
  labs(
    x = NULL,
    y = NULL
  )
```

# Application exercise {.smaller}

## `{r} todays_ae`

::: appex
-   Go to your ae project in RStudio.

-   If you haven't yet done so, make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you haven't yet done so, click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Work through the application exercise in class, and render, commit, and push your edits.
:::
