---
title: "Final Thoughts: Statistical inference"
subtitle: "Lecture 23"
date: "2025-6-18"
format: 
  revealjs:
    output-file: 23-inference-recap-slides.html
    pdf-separate-fragments: true
auto-stretch: false
---

# Recap: statistical inference

```{r}
#| label: load-packages
#| message: false
#| echo: false
library(tidyverse)
library(tidymodels)
library(openintro)
set.seed(123)
counties_sample <- county_2019 |>
  sample_n(25)

new_loans <- loans_full_schema |>
  drop_na(annual_income, total_credit_utilized) |>
  filter(log(annual_income) > 0) |>
  filter(log(total_credit_utilized) > 0)
```

## Point estimation: what's your best guess? {.smaller}

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

## Point estimation: what's your best guess? {.smaller}

```{r}
linear_reg() |>
  fit(mpg ~ wt, data = mtcars) |>
  tidy()
```

## Sampling variability

Different data \>\> Different estimate.
But how different?

-   **Reassuring**: recompute your estimate on other data and get basically the same answer.
    Sampling variation is low.
    Uncertainty is low.
    Results are more reliable;

-   **Concerning**: recompute your estimate on new data and get a wildly different answer.
    Sampling variation is high.
    Uncertainty is high.
    Results are less reliable.

How do we operationalize this?

## The bootstrap

Construct different, alternative datasets by sampling *with replacement* from your original data.
Each bootstrap sample gives you a different estimate, and you can see how vigorously they wiggle around:

```{r}
#| echo: false
#| message: false
#| warning: false
df_boot_samples_100 <- new_loans |>
  mutate(
    log_cred = log(total_credit_utilized),
    log_inc = log(annual_income)
  ) |> 
  slice(sample(1:9000, 100)) |>
  specify(log_cred ~ log_inc) |>
  generate(reps = 200, type = "bootstrap")
```

::: columns
::: {.column width="49%"}
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 1
p_df_boot_samples_100 <- ggplot(df_boot_samples_100, aes(x = log_inc, y = log_cred, group = replicate)) +
  geom_line(stat = "smooth", method = "lm", se = FALSE, alpha = 0.05) +
  labs(
    x = "Annual income (log $)",
    y = "Credit utilization (log $)",
  ) +
  #xlim(7, 15) + 
  #ylim(0, 15) + 
  theme(title = element_text(size = 20, face = "bold"))

p_df_boot_samples_100
```
:::

::: {.column width="49%"}
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-asp: 1
df_boot_samples_100_fit <- df_boot_samples_100 |>
  fit()

df_boot_samples_100_hist <- ggplot(df_boot_samples_100_fit |> filter(term == "log_inc"), aes(x = estimate)) +
  geom_histogram(color = "white") +
  #geom_vline(xintercept = slope, color = "#8F2D56", linewidth = 1) +
  labs(x = "Slope estimate", y = "Count",
       title = "Histogram of alternative estimates") + 
  theme(title = element_text(size = 20, face = "bold"))

df_boot_samples_100_hist
```
:::
:::

## Interval estimation

Use the bootstrap distribution to construct a range of likely values for the unknown parameter:

```{r}
#| echo: false

df_fit <- linear_reg() |>
  fit(price ~ area, data = duke_forest)

intercept <- tidy(df_fit) |> filter(term == "(Intercept)") |> pull(estimate) |> round()
slope <- tidy(df_fit) |> filter(term == "area") |> pull(estimate) |> round()

set.seed(119)

df_boot_samples_100 <- duke_forest |>
  specify(price ~ area) |>
  generate(reps = 100, type = "bootstrap")

df_boot_samples_100_fit <- df_boot_samples_100 |>
  fit()

df_boot_samples_100_hist <- ggplot(df_boot_samples_100_fit |> filter(term == "area"), aes(x = estimate)) +
  geom_histogram(binwidth = 10, color = "white") +
  geom_vline(xintercept = slope, color = "#8F2D56", linewidth = 1) +
  labs(x = "Slope", y = "Count",
       title = "Slopes of 100 bootstrap samples") +
  scale_x_continuous(labels = label_dollar())

lower <- df_boot_samples_100_fit |>
  ungroup() |>
  filter(term == "area") |>
  summarise(quantile(estimate, 0.025)) |>
  pull()

upper <- df_boot_samples_100_fit |>
  ungroup() |>
  filter(term == "area") |>
  summarise(quantile(estimate, 0.975)) |>
  pull()

df_boot_samples_100_hist +
  geom_vline(xintercept = lower, color = "steelblue", linewidth = 1, linetype = "dashed") +
  geom_vline(xintercept = upper, color = "steelblue", linewidth = 1, linetype = "dashed")
```

We use quantiles (think IQR), but there are other ways.

## Hypothesis testing {.smaller}

::: incremental
-   Two competing claims about $\beta_1$: $$
    \begin{aligned}
    H_0&: \beta_1=0\quad(\text{nothing going on})\\
    H_A&: \beta_1\neq0\quad(\text{something going on})
    \end{aligned}
    $$

-   Do the data strongly favor one or the other?

-   How can we quantify this? p-values!
:::

## *p*-value {.scrollable}

*p*-value is the fraction of the histogram area shaded red:

```{r}
#| echo: false
#| message: false
#| warning: false
set.seed(3)
x <- rnorm(100, mean = 0.1, sd = 2)
df <- tibble(x = x)
obs_stat <- df |>
  specify(response = x) |>
  calculate(stat = "mean")
null_dist <- df |>
  specify(response = x) |>
  hypothesize(null = "point", mu = 0) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean")


visualize(null_dist) + 
  shade_p_value(obs_stat = obs_stat, direction = "two-sided")
```


## Discernibility level {.scrollable}

::: incremental
-   How do we decide if the *p*-value is big enough or small enough?

-   Pick a threshold $\alpha\in[0,\,1]$ called the *discernibility level*:

    -   If $p\text{-value} < \alpha$, reject null and accept alternative;
    -   If $p\text{-value} \geq \alpha$, fail to reject null;

-   Standard choices: $\alpha=0.01, 0.05, 0.1, 0.15$.
:::

## AE 17 Recap

::: incremental
-   **Goal**: Predicting mean household income using the percent of a county that is uninsured

-   We did inference with `counties_sample`, a small sample of 25 counties in the US
:::

## Point Estimate {.smaller}

```{r}
observed_fit <- counties_sample |>
  specify(mean_household_income ~ uninsured) |>
  fit()

observed_fit
```

## Generating the null distribution {.smaller}

```{r}
#| warning: false
#| message: false
#| fig-asp: 0.3
null_dist <- counties_sample |>
  specify(mean_household_income ~ uninsured) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()

null_dist |> 
  filter(term == "uninsured") |>
  ggplot(aes(x = estimate)) + 
  geom_histogram()
```

## How does this work? {.smaller}

```{r}
null_dist <- counties_sample |>
  specify(mean_household_income ~ uninsured) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()

null_dist
```

## How does this work? {.smaller}

```{r}
#| eval: false
#| code-line-numbers: "2"
null_dist <- counties_sample |>
  specify(mean_household_income ~ uninsured) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()
```

## How does this work? {.smaller}

```{r}
#| eval: false
#| code-line-numbers: "3"
null_dist <- counties_sample |>
  specify(mean_household_income ~ uninsured) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()
```

## How does this work? {.smaller}

```{r}
#| eval: false
#| code-line-numbers: "4"
null_dist <- counties_sample |>
  specify(mean_household_income ~ uninsured) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()
```

## What is happening? {.smaller}

::: incremental
-   Recall: Bootstrapping for confidence intervals

    -   We created a bunch of new data sets by **resampling** from our original data set

    -   We computed an estimate for each new data set

-   Here:

    -   We are creating a bunch of new data sets by **permuting** our original data set and computing an estimate for each

    -   Instead of attempting to mimic samples from the population data (like boostrapping), we are attempting to mimic samples from the *null*
:::

## So what does "permuting" do? {.smaller}

We are simulating what the world would look like *if* the **null hypothesis were true**: what if `mean_household_income` and `uninsured` are unrelated?

::: incremental
-   We **randomly shuffle** the values of the response variable: `mean_household_income`
-   We **leave** the explanatory variable (`uninsured`) unchanged
-   We do this many times (e.g. `reps = 1000`)
-   Each time we calculate the slope from the new fake dataset
:::

## Small permutation example: {.smaller}

::: columns
::: {.column width="50%"}
**Original:**

| Uninsured (%) | Mean Income (\$k) |
|---------------|-------------------|
| 5             | 70                |
| 8             | 65                |
| 6             | 68                |
| 7             | 66                |
| 9             | 64                |
:::

::: {.column width="50%"}
:::
:::

## Small permutation example: {.smaller}

::: columns
::: {.column width="50%"}
**Original:**

| Uninsured (%) | Mean Income (\$k) |
|---------------|-------------------|
| 5             | 70                |
| 8             | 65                |
| 6             | 68                |
| 7             | 66                |
| 9             | 64                |
:::

::: {.column width="50%"}
**Permutation 1:**

| Uninsured (%) | Mean Income (\$k) |
|---------------|-------------------|
| 8             | 70                |
| 5             | 65                |
| 7             | 68                |
| 9             | 66                |
| 6             | 64                |

:::
:::

## Small permutation example: {.smaller}

::: columns
::: {.column width="50%"}
**Original:**

| Uninsured (%) | Mean Income (\$k) |
|---------------|-------------------|
| 5             | 70                |
| 8             | 65                |
| 6             | 68                |
| 7             | 66                |
| 9             | 64                |
:::

::: {.column width="50%"}
**Permutation 2:**

| Uninsured (%) | Mean Income (\$k) |
|---------------|-------------------|
| 5             | 70                |
| 7             | 65                |
| 9             | 68                |
| 8             | 66                |
| 6             | 64                |

:::
:::

. . . 

Do this `reps` times; make a histogram of slope estimates!

## Null distribution {.smaller}

The null distribution shows the plot of 
```{r}
#| echo: false
#| message: false
null_dist |> 
  filter(term == "uninsured") |>
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 30)+
  labs(title = "Null distribution of slope",
       x = "Slope estimate under null",
       y = "Count")
```



## Then what? {.smaller}

Compare the slope from the original data set to our null distribution!

```{r}
#| echo: false
#| message: false
null_dist |> 
  filter(term == "uninsured") |> 
  ggplot(aes(x = estimate)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = observed_fit |> filter(term == "uninsured") |> pull(estimate),
             color = "red", linewidth = 2) +
  labs(title = "Null distribution of slope (uninsured)",
       x = "Slope estimate under null",
       y = "Count")
```

## Better code {.smaller}

```{r}
#| label: shade-p-value
#| eval: true
visualize(null_dist) +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```


## The p-value {.smaller}

```{r}
null_dist |>
  get_p_value(obs_stat = observed_fit, direction = "two-sided")
```


## Recap: Hypothesis testing {.smaller}

::: incremental
-   Think hypothetically: if the null hypothesis were in fact true, would my results be out of the ordinary?

    -   if no, then the null could be true;
    -   if yes, then the null might be bogus;

-   My results represent the reality of actual data.
    If they conflict with the null, then you throw out the null and stick with reality;

-   How do we *quantify* "would my results be out of the ordinary"?
:::

## Idea: Hypothesis test for a difference in means {.smaller}

What if I want to test if the means of two groups are actually different from each other? 

:::incremental
- Is the mean household income different in counties that vote red vs. blue?
- Do athletes who train over a certain number of hours a day have better performance (measured by some metric) than those who train under that number?
- Do babies born to mothers who smoke have lower body weights than non-smoking mothers?

:::

## Example: smoking/non-smoking mothers {.smaller}

A random sample of US births from 2014:

```{r}
#| attr-output: "style='font-size: 0.6em'"

births14 <- births14 |> drop_na(weight, habit) 
births14
```

Do babies born to mothers who smoke have lower body weights than non-smoking mothers?

## Visualize {.smaller}

```{r}
births14 |>
  ggplot(aes(x = weight, y = habit)) + 
  geom_boxplot(alpha = 0.5) + 
  labs(
    x = "Birth Weight (lbs)",
    y = "Mother smoking status"
  )
```

## Estimate {.smaller}

```{r}
linear_reg() |>
  fit(weight ~ habit, data = births14) |>
  tidy()
```

$$
\widehat{weight}=7.27-0.59\times \texttt{smoker}.
$$

## Remember... {.smaller}

$$
\widehat{weight}=7.27-0.59\times \texttt{smoker}.
$$

::: incremental
-   `habit` is categorical with two levels: `smoker` and `non-smoker`.

-   The baseline is `non-smoker`, with $$
    \widehat{weight} = 7.27;
    $$

-   The smoking group has `smoker = 1`, so $$
    \widehat{weight} = 7.27 - 0.59 = 6.68.
    $$
:::

## Those were not magic numbers {.smaller}

```{r}
births14 |>
  group_by(habit) |>
  summarize(
    mean_weight = mean(weight)
  )
```

. . . 


::: callout-note
## Don't forget!

In a simple linear regression with a numerical response and a binary predictor, the slope estimate is the difference in the average response between the two groups.
:::
## Confidence interval {.smaller}

```{r}
#| message: false
#| warning: false
#| echo: false

set.seed(1234)

boot_fits <- births14 |>
  specify(weight ~ habit) |>
  generate(reps = 1000, type = "bootstrap") |>
  fit()

observed_fit <- births14 |>
  specify(weight ~ habit) |>
  fit()

ci_95 <- get_confidence_interval(
  boot_fits,
  point_estimate = observed_fit,
  level = 0.95,
  type = "percentile"
)

boot_fits |>
  filter(term == "habitsmoker") |>
  ggplot(aes(x = estimate)) +
  geom_histogram() +
  geom_vline(xintercept = observed_fit |> filter(term == "habitsmoker") |> pull(estimate),
             color = "#8F2D56", linewidth = 1) + 
  geom_vline(xintercept = ci_95 |> filter(term == "habitsmoker") |> pull(lower_ci), 
             color = "steelblue", size = 1, linetype = "dashed") + 
  geom_vline(xintercept = ci_95 |> filter(term == "habitsmoker") |> pull(upper_ci),
             color = "steelblue", size = 1, linetype = "dashed") + 
  labs(
    title = "Boostrap distribution",
    x = "Estimated difference (Smoker - Non-smoker)"
  )
```

We are 95% confident that the true mean difference is between -0.919 and -0.301. This does not contain 0!!!

## Competing claims {.smaller}

Are the data sufficiently informative to tell the difference between these two possibilities?

$$
\begin{aligned}
H_0&: \beta_1=0 && (\text{no effect})\\
H_0&: \beta_1\neq0 && (\text{some effect})\\
\end{aligned}
$$

:::incremental
-   $\beta_1$ is the difference in mean response between the smoking and non-smoking group;
-   the alternative does not specify if the effect is positive or negative;
-   the alternative does not specify that the effect is "large".
:::

## Hypothesis test {.smaller}

```{r}
#| message: false
#| warning: false
#| echo: false

set.seed(1234)
null_dist <- births14 |>
  specify(weight ~ habit) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()


visualize(null_dist) +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```

. . .

The p-value is basically zero.

## What did we learn? {.smaller}

::: incremental
-   Evidence is sufficient to reject the null: smoking during pregnancy is *associated* with a difference in birth weight;

-   Is smoking *causing* birth weight to change?

-   The data are observational, and there are many unmeasured factors that could bear on both `habit` and `weight`.

-   Maybe there is a causal relationship, and maybe it can be teased out with data like these, but it is not so straightforward
:::

. . .

> Just because there is a statistically discernible association does not mean there is a causal relationship.


## So, when can we draw causal conclusions? {.smaller}

::: incremental
-   Best hope: run a controlled, randomized experiment:

    -   get a big, representative, random sample from target population;
    -   randomly divide subjects into treatment and control;
    -   difference in outcomes was *caused* by the treatment alone;
    -   but this may be too costly or unethical.

-   Observational data are very challenging:

    -   you couldn't control gosh darned thing;
    -   different outcomes could be caused by confounding factors, which you may or may not have measured.
:::


## A proper, controlled experiment {.smaller}

Does treatment using embryonic stem cells (ESCs) help improve heart function following a heart attack?
Each sheep in the study was randomly assigned to the ESC or control group, and the change in their hearts' pumping capacity was measured in the study.
A positive value corresponds to increased pumping capacity, which generally suggests a stronger recovery.

. . .

```{r}
#| attr-output: "style='font-size: 0.5em'"
glimpse(stem_cell)
```

## Visualize {.smaller}

```{r}
stem_cell <- stem_cell |>
  mutate(change = after - before)

ggplot(stem_cell, aes(x = change, y = trmt)) + 
  geom_boxplot()
```

## Model fit {.smaller}

```{r}
linear_reg() |>
  fit(change ~ trmt, data = stem_cell) |>
  tidy()
```

$$
\widehat{change}=-4.33+7.83\times \texttt{esc}
$$


## Remember... {.smaller}

$$
\widehat{change}=-4.33+7.83\times \texttt{esc}
$$

::: incremental
-   `trmt` is categorical with two levels: `ctrl` (0) and `esc` (1).

-   The control group has `esc = 0`, so $$
    \widehat{change} = -4.33;
    $$

-   The treatment group has `esc = 1`, so $$
    \widehat{change} = -4.33+7.83=3.5.
    $$
:::

## Again, those were not magic numbers!! {.smaller}
 
```{r}
stem_cell |>
  group_by(trmt) |>
  summarize(
    mean_change = mean(change)
  )
```


## Confidence interval {.smaller}

```{r}
#| message: false
#| warning: false
#| echo: false

set.seed(1234)

boot_fits <- stem_cell |>
  specify(change ~ trmt) |>
  generate(reps = 1000, type = "bootstrap") |>
  fit()

observed_fit <- stem_cell |>
  specify(change ~ trmt) |>
  fit()

ci_95 <- get_confidence_interval(
  boot_fits,
  point_estimate = observed_fit,
  level = 0.95,
  type = "percentile"
)

boot_fits |>
  filter(term == "trmtesc") |>
  ggplot(aes(x = estimate)) +
  geom_histogram() +
  geom_vline(xintercept = observed_fit |> filter(term == "trmtesc") |> pull(estimate),
             color = "#8F2D56", linewidth = 1) + 
  geom_vline(xintercept = ci_95 |> filter(term == "trmtesc") |> pull(lower_ci), 
             color = "steelblue", size = 1, linetype = "dashed") + 
  geom_vline(xintercept = ci_95 |> filter(term == "trmtesc") |> pull(upper_ci),
             color = "steelblue", size = 1, linetype = "dashed") + 
  labs(
    title = "Boostrap distribution",
    x = "Estimated difference (Treament - Control)"
  )
```

We are 95% confident that the true mean difference is between 4.34 and 11.7.

## Competing claims: again! {.smaller}

Are the data sufficiently informative to tell the difference between these two possibilities?

$$
\begin{aligned}
H_0&: \beta_1=0 && (\text{no effect})\\
H_0&: \beta_1\neq0 && (\text{some effect})\\
\end{aligned}
$$

-   $\beta_1$ is the difference in mean response between the treatment and control group;
-   the alternative does not specify if the effect is positive or negative;
-   the alternative does not specify that the effect is "large".

## Hypothesis test {.smaller}

```{r}
#| message: false
#| warning: false
#| echo: false

set.seed(1234)
null_dist <- stem_cell |>
  specify(change ~ trmt) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()


visualize(null_dist) +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```

. . .

The p-value is basically zero.

## What did we learn? {.smaller}

::: incremental
-   In a hypothetical world where the treatment has no effect, it would be very unlikely to get results like the ones we did;

-   Evidence is sufficient to *reject the null*.
    There is *some* effect;

-   If you are convinced by the experimental design, the effect is *causal*;

-   Is the effect large and meaningful and something you should base major decisions on?
    Statistics cannot answer that question!
:::


# AE !

Today's AE: Inference Overview

## Announcements {.smaller}

- I am filling in for Mary's OH today!!

- Do your second peer feedback forms!!

- Come to lab on Friday!!!