---
title: "AE 02: Penguin peekaboo"
author: Your Name
---

For this application exercise, we'll use the **tidyverse** and **palmerpenguins** packages.

```{r}
#| label: load-packages
library(tidyverse)
library(palmerpenguins)
```

We will again be working with the `penguins` dataset for this application exercise.
Let's `glimpse()` at the dataset to remind ourselves of what data we have at our fingertips.

-   **Your turn:** Replace `# add code here` with the code for "glimpse"ing at the data `penguins` data frame – `glimpse(penguins)`.

```{r}
#| label: glimpse-penguins
# add code here
```

# Task 1

In the code chunk below, plot a histogram of the body mass of the penguins. Let's also *facet* by island:

::: callout-note
All plots created in this course **must** be properly labeled (i.e., human readable labels)!
This means both x- and y-axes (when appropriate) and an informative but concise title!
Double check your rendered document to ensure your plot labels are fully visible. 
:::

```{r}
#| label: mass-hist

# add code here
```

# Task 2

In the code chunk below, plot a scatterplot with body mass on the horizontal and flipper length on the vertical, colored by species:

```{r}
#| label: mass-flipper-scatter

# add code here
```
