---
title: "AE 02: Penguin peekaboo"
author: Your Name
---

For this application exercise, we'll use the **tidyverse** and **palmerpenguins** packages.

```{r}
#| label: load-packages
library(tidyverse)
library(palmerpenguins)
```

The dataset we will visualize is called `penguins`.
Let's `glimpse()` at it.

-   **Your turn:** Replace `# add code here` with the code for "glimpse"ing at the data `penguins` data frame – `glimpse(penguins)`. Render the document and view the output.

```{r}
#| label: glimpse-penguins
# add code here
```

-   **Demo:** First, replace the blank below with the number of rows in the `penguins` data frame based on the output of the chunk below. Then, replace it with "inline code" and render again.

```{r}
#| label: nrow-penguins
nrow(penguins)
```

There are \_\_\_ penguins in the `penguins` data frame.

# Task 1

In the code chunk below, plot a histogram of the body mass of the penguins:

```{r}
#| label: mass-hist

# add code here
```

# Task 2

In the code chunk below, plot a scatterplot with body mass on the horizontal and flipper length on the vertical:

```{r}
#| label: mass-flipper-scatter

# add code here
```
