---
title: "AE 07: Joining"
format:
  html:
    self-contained: true
---

## Goal

Let's suppose we were interested in examining the total populations of continents.
However, these are the two data sets we have available:

-   world-pop-22.csv: contains data from The World Bank with countries' populations

-   continents.csv: contains data mapping countries to continents

To do analysis of population by continent, we need to join!

#### Load packages and read in data

```{r}
library(tidyverse)
population <- read_csv("data/world-pop-2022.csv")
continents <- read_csv("data/continents.csv")
```

#### Take a look at the data:

```{r}
population
continents
```

### Question 1: Join Concept

We want to know what continent *all* of the variables in the `population` data frame are in.

-   What type of join should we use?

-   Which variable in each data frame should we use?

### Question 2: Implement the Join

Join the two data frames and name assign the joined data frame to a new data frame `population_continents` .

```{r}

```

How does that look?
Take a look at your new data frame!

### Question 3: What went wrong?

It might not be obvious, but something is a little weird about this.
Go ahead and `filter` the resulting data frame to see if any of the continent values are `NA`.

```{r}

```

There are!
This means that there were no rows in the `continents` data frame with those countries.
That seems a little weird.
Take a scroll through the continents data frame.

***Do you see what the cause of this is??***

### Question 4: Let's fix this!

So, countries have to be spelled the exact same way in each data set.
I'm going to show you code that renames the missing countries in the `population` data set to match the spelling in `contintents`.
Then, we will re-run the join: we are no longer missing these values!

```{r}
population_new_spellings <- population |>
  mutate(country = case_when(
    country == "Congo, Dem. Rep." ~ "Democratic Republic of Congo",
    country == "Congo, Rep." ~ "Congo",
    country == "Hong Kong SAR, China" ~ "Hong Kong",
    country == "Korea, Dem. People's Rep." ~ "North Korea",
    country == "Korea, Rep." ~ "South Korea",
    country == "Kyrgyz Republic" ~ "Kyrgyzstan",
    .default = country
    )
  ) 
```

```{r}
population_continent_new = population_new_spellings |>
  left_join(continents, by = join_by(country == entity))
```

Are there any NAs left?
Run the code to see!

```{r}
population_continent_new |>
  filter(is.na(continent))
```

## The Bigger Picture

How does what we just did fit into the bigger picture of the class?

-   You are almost never going to be joining data sets just for fun!
    Joining data sets is a way to enable investigating more interesting patterns in data.

-   Now, you could create plots or compute summary statistics to investigate the relationship between continent and total population.

***For extra practice,*** try making a bar plot of total population in each country.

***To see something new,*** try making a 'lollipop chart' of population by country.
