---
title: "AE 16: Forest classification 🌳🌲"
subtitle: "Suggested answers"
format:
  html:
    self-contained: true
---

In this application exercise, we will

-   Split our data into testing and training
-   Fit logistic regression regression models to testing data to classify outcomes
-   Evaluate performance of models on testing data

We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively, and the **forested** package for the data.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(forested)
```

Remember from the lecture that the `forested` dataset contains information on whether a plot is forested (`Yes`) or not (`No`) as well as numerical and categorical features of that plot.

```{r}
#| label: glimpse-forested
glimpse(forested)
```

# Spending your data

Split your data into testing and training in a reproducible manner and display the split object.

```{r}
#| label: train-test-split
set.seed(253)

forest_split <- initial_split(forested)
forest_split
```

What percent of the original `forested` data is allocated to training and what percent to testing?
Compare your response to your neighbor's.
Are the percentages roughly consistent?
What determines this in the `initial_split()`?
How would the code need to be updated to allocate 80% of the data to training and the remaining 20% to testing?

\~75% of the data is allocated to training, by default.

```{r}
#| label: train-test-percentages
5330/7107
```

To allocate 80% to training, we need to add the argument `prop = .8` to our `initial_split()` function.

```{r}
#| label: train-test-80-20
forested_split_2 <- initial_split(forested, prop = .8)
forested_split_2
```

Let's stick with the default split and save our testing and training data.

```{r}
#| label: train-test-save
forest_train <- training(forest_split)
forest_test <- testing(forest_split)
```

# Exploratory data analysis

Create a visualization that explores the relationship between the outcome, one numerical predictor, and one categorical predictor.
Then, describe, in a few sentences, what the visualization shows about the relationship between these variables.

**Note:** Pay attention to which dataset you use for your exploration.

```{r}
#| label: eda-plot
#| warning: false
#| message: false
ggplot(forest_train, aes(x = temp_annual_mean, fill = land_type)) +
  geom_histogram(alpha = 0.7) +
  facet_wrap(~ forested, ncol = 1)
```

We observe that that the majority of plots classified as not forested are of land type "Barren" or "Non-tree vegetation", not "Tree".
The distribution of `temp_annual_mean` is centered at a higher value for non-forested plots than for forested, and it's more concentrated around it's center as well.

# Model 1: Custom choice of predictors

## Fit

Fit a model for classifying plots as forested or not based on a subset of predictors of your choice.
Name the model `forested_custom_fit` and display a tidy output of the model.

```{r}
#| label: forested-custom-fit
forested_custom_fit <- logistic_reg() |>
  fit(forested ~ precip_annual + elevation + temp_annual_mean, data = forest_train)

tidy(forested_custom_fit)
```

## Interpret

Interpret the intercept (with respect to $\widehat{p}$) in the context of these data.
Then, interpret one slope coefficient of your choice (with respect to the odds), again in the context of these data.

Intercept:

For a plot of land with 0 annual precipitation, 0 elevation, and an average annual temperature 0, we predict the probability that the plot is not forested is 0.00181323.

All else equal, we expect that for a one unit increase in annual precipitation, the predicted odds that a plot of land is not forested will decrease by a multiplicative factor of 0.9982989, on average.

```{r}
#| label: forested-interpret
exp(-6.310830752) / (1 + exp(-6.310830752	))

levels(forest_train$forested)

exp(-0.001702569)
```

## Predict

Predict for the testing data using this model.

```{r}
#| label: forested-custom-aug
forest_aug <- augment(forested_custom_fit, new_data = forest_test)

forest_aug
```

## Evaluate

Calculate the error rates (and their complements) for the testing data using this model.

```{r}
#| label: forested-custom-eval
forest_aug |>
  count(forested, .pred_class) |>
  group_by(forested) |>
  mutate(prop = n / sum(n),
         decision = case_when(
           forested == "Yes" & .pred_class == "Yes" ~ "True Positive",
           forested == "Yes" & .pred_class == "No" ~ "False Negative",
           forested == "No" & .pred_class == "Yes" ~ "False Positive",
           forested == "No" & .pred_class == "No" ~ "True Negative"
         ))
```

Another commonly used display of this information is a confusion matrix.
Create this using the `conf_mat()` function.
You will need to review the documentation for the function to determine how to use it.

```{r}
#| label: conf-mat-custom
conf_mat(forest_aug,
         truth = forested,
         estimate = .pred_class)
```

## Sensitivity, specificity, ROC curve

Calculate sensitivity and specificity and draw the ROC curve.

```{r}
#| label: forested-custom-roc
forest_custom_roc <- roc_curve(forest_aug,
                               truth = forested,
                               .pred_Yes,
                               event_level = "first")

forest_custom_roc

ggplot(forest_custom_roc, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = "dashed")
```

# Model 2: All predictors

## Fit

Fit a model for classifying plots as forested or not based on all predictors available.
Name the model `forested_full_fit` and display a tidy output of the model.

```{r}
#| label: forested-full-fit
forested_full_fit <- logistic_reg() |>
  fit(forested ~ ., data = forest_train)

tidy(forested_full_fit)
```

## Predict

Predict for the testing data using this model.

```{r}
#| label: forested-full-aug
forest_full_aug <- augment(forested_full_fit, new_data = forest_test)
forest_full_aug
```

## Evaluate

Calculate the error rates (and their complements) for the testing data using this model.

```{r}
#| label: forested-full-eval
forest_full_aug |>
  count(forested, .pred_class) |>
  group_by(forested) |>
  mutate(prop = n / sum(n),
         decision = case_when(
           forested == "Yes" & .pred_class == "Yes" ~ "True Positive",
           forested == "Yes" & .pred_class == "No" ~ "False Negative",
           forested == "No" & .pred_class == "Yes" ~ "False Positive",
           forested == "No" & .pred_class == "No" ~ "True Negative"
         ))
```

## Sensitivity, specificity, ROC curve

Calculate sensitivity and specificity and draw the ROC curve.

```{r}
#| label: forested-full-roc
forest_full_roc <- roc_curve(forest_full_aug,
                               truth = forested,
                               .pred_Yes,
                               event_level = "first")

ggplot(forest_full_roc, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = "dashed")
```

# Model 1 vs. Model 2

Plot both ROC curves and articulate how you would use them to compare these models.

```{r}
#| label: compare
forest_custom_roc <- forest_custom_roc |>
  mutate(model = "custom")
forest_custom_roc

forest_full_roc <- forest_full_roc |>
  mutate(model = "full")

forest_roc <- rbind(forest_custom_roc, forest_full_roc)
forest_roc

ggplot(forest_roc, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_path() +
  geom_abline(lty = "dashed")
```

We would choose to favor the model with the larger AUC; in this case, this is the full model.

```{r}
forest_full_auc <- roc_auc(forest_full_aug,
                               truth = forested,
                               .pred_Yes,
                               event_level = "first")
forest_full_auc

forest_custom_auc <- roc_auc(forest_aug,
                               truth = forested,
                               .pred_Yes,
                               event_level = "first")
forest_custom_auc
```
