---
title: "AE 23: Pokemon Inference"
---

Today we’ll explore the question “Are taller Pokémon speedier than shorter Pokémon?”

# Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
```

# Data

The data for this application exercise was originally gathered using the [**Complete Pokémon Dataset**](https://www.kaggle.com/datasets/rounakbanik/pokemon) from Kaggle. 

First, let's load the data:

```{r}
#| label: load-data
#| message: false

pokemon <- read_csv("data/pokemon.csv")
```

To keep things simple, we'll work with a subset of the data, the generation 4 Pokémon.

```{r}
pokemon <- pokemon |>
  filter(generation == 4)

glimpse(pokemon)
```

We are interested in whether a Pokémon's log height (measured in meters) can predict its speed (how fast it is).

# Visualize

a. Review! Create a new variable for the Pokémon's log height.


b. **Your turn**: Plot the data and the line of best fit.

```{r}
#| message: false
# add code here
```

# Point Estimation

a.  **Your turn**: Fit the linear model to these data:

```{r}
#| message: false
# add code here
```

::: callout-note 
This gives the exact same numbers that you get if you use `linear_reg() |> fit()`, but we need this new syntax because it plays nice with the tools we have for confidence intervals and hypothesis tests. I know, I hate it too, but it's the way it is.
:::

b.  **Your turn**: Typeset the equation for the model fit:

$$
YEEHAW
$$

c.  **Your turn**: Interpret the slope and the intercept estimates:


# Hypothesis Testing

Let's consider the hypotheses:

$$
H_0:\beta_1=0\quad vs\quad H_A: \beta_1\neq 0.
$$ The null hypothesis corresponds to the claim that log_height and speed are uncorrelated.

a.  Simulate and plot the null distribution for the slope:

```{r}
#| message: false
set.seed(24601)
# add code here
```

b.  Add a vertical line to your plot indicating the point estimate of the slope from your original data analysis:

```{r}
# add code here
```

c.  Compute the $p$-value for this test and interpret it:

```{r}
#| warning: false
# add code here
```

# Interval Estimation

a.  **Demo**: Using seed `24601`, generate `500` bootstrap samples, and store them in a new data frame called `bstrap_samples`.

```{r}
set.seed(24601)
# add code here
```

b.  **Demo**: Fit a linear model to each of these bootstrap samples and store the estimates in a new data framed called `bstrap_fits`.

```{r}
# add code here
```

c.  **Your turn**: Use `linear_reg() |> fit(...)` to fit a linear model to bootstrap sample number *347*, and verify that you get the same estimates as the ones contained in `bstrap_fits`.

```{r}
# add code here
```

::: callout-note 
The only point I'm making here is that this new bootstrap code is not performing a fundamentally new task. It's performing an old task (fitting the linear model), but it's repeating it A LOT. So the numbers you get here are not mysterious. They're numbers you already know how to compute.
:::

d.  **Demo**: Compute 95% confidence intervals for the slope and the intercept using the `get_confidence_interval` command.

```{r}
# add code here
```

e.  **Your turn**: Verify that you get the same numbers when you manually calculate the quantiles of the slope estimates using `summarize` and `quantile`. *Pay attention to the grouping*.

```{r}
# add code here
```

::: callout-note 
Same point as before. There's no magic here. `get_confidence_interval` is just a convenient way of doing something that you already knew how to do.s
:::

f. **BONUS**: You can visualize the confidence interval:

```{r}
# add code here
```


