---
title: "AE 18: FIFA Inference ⚽"
subtitle: "Suggested answers"
format:
  html:
    self-contained: true
---

Today we’ll explore the relationship between a FIFA member country's share of the world population and it's share of the global World Cup TV audience.

# Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
```

# Data

In this exercise, we will work with data from FiveThirtyEight’s story *How To Break FIFA*. The dataset contains information on FIFA member countries, including each country’s confederation, share of the global population, share of World Cup TV audience, and GDP-weighted audience share.

The dataset, called `fifa_countries_audience.csv`, can be found in the `data` folder.

First, let's load the data:

```{r}
#| label: load-data
#| message: false
fifa <- read_csv("data/fifa_countries_audience.csv")

glimpse(fifa)
```

## Data Dictionary

| Variable | Description |
|-----------------|-------------------------------------------------------|
| `country` | FIFA member country |
| `confederation` | Confederation to which the country belongs |
| `population_share` | Country’s share of the global population, recorded as a percentage |
| `tv_audience_share` | Country’s share of the global World Cup TV audience, recorded as a percentage |
| `gdp_weighted_share` | Country’s GDP-weighted audience share, recorded as a percentage |

# ABV (Always Be Visualizing)

a.  **Your turn**: First, let's get rid of two outliers. Create a new data frame called `fifa_clean` that contains only those member countries whose share of the global population is below 10%.

```{r}
fifa_clean <- fifa |>
  filter(population_share < 10)
```

b.  **Your turn**: Create a visualization to explore the relationship between our variables of interest. Let's consider `population_share` as our explanatory variable and `tv_audience_share` as our outcome. Include a linear trendline.

```{r}
#| message: false
ggplot(fifa_clean, aes(x = population_share, y = tv_audience_share)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Now let’s imagine we only had a tiny subset of these data to work with:

```{r}
#| eval: true
set.seed(847)
baby_fifa <- fifa_clean |>
  slice(sample(1:nrow(fifa_clean), 25))
glimpse(baby_fifa)
```

c.  Plot the baby thing, again adding a linear trendline:

```{r}
ggplot(baby_fifa, aes(x = population_share, y = tv_audience_share)) +
  geom_point() +
  geom_smooth(method = "lm")
```

# Inference with the small dataset

a.  **Your turn**: Obtain the point estimate $b_1$ from the baby data.

```{r}
#| message: false
observed_fit <- baby_fifa |>
  specify(tv_audience_share ~ population_share) |>
  fit()

observed_fit
```

::: callout-note
This gives the exact same numbers that you get if you use `linear_reg() |> fit()`, but we need this new syntax because it plays nice with the tools we have for confidence intervals and hypothesis tests.
I know, I hate it too, but it's the way it is.
:::

b.  **Your turn**: Typeset the equation for the model fit:

$$
\widehat{tv~audience~share} = -0.09415766   + 2.45268107 \times population~share
$$

c.  **Your turn**: Interpret the slope and the intercept estimates:

We predict that a FIFA member country with 0% share of the global population has a -9.4% share of the global World Cup TV audience, on average.

We would expect that for every percentage point increase in a FIFA member country's share of the global population, the country's share of the global World Cup TV audience increases by 2.45 percentage points, on average.

# Hypothesis Testing

Let's consider the hypotheses:

$$
H_0:\beta_1=0
$$ $$
H_A: \beta_1\neq 0.
$$ The null hypothesis corresponds to the claim that a FIFA member country's share of the global population and its share of the global World Cup TV audience are unrelated / uncorrelated.

a.  Simulate and plot the null distribution for the slope:

```{r}
#| message: false
set.seed(847)

null_dist <- baby_fifa |>
  specify(tv_audience_share ~ population_share) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()

null_dist

null_dist |>
  filter(term == "population_share") |>
  ggplot(aes(x = estimate)) +
  geom_histogram()
```

b.  Where does our actual point estimate fall under the null distribution? Add a vertical line corresponding to our point estimate and *shade* the region corresponding to the $p$-value.

```{r}
visualize(null_dist) +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```

c.  Compute the $p$-value for this test and interpret it:

```{r}
#| warning: false
null_dist |>
  get_p_value(obs_stat = observed_fit, direction = "two-sided")
```

Reject the null hypothesis.
There is sufficient evidence for some relationship between a FIFA member country's share of the global population and its share of the global World Cup TV audience.

# Interval Estimation

a.  **Demo**: Generate `500` bootstrap samples, and store them in a new data frame called `bstrap_samples`.

```{r}
set.seed(847)

bstrap_samples <- baby_fifa |>
  specify(tv_audience_share ~ population_share) |>
  generate(reps = 500, type = "bootstrap")
```

b.  **Demo**: Fit a linear model to each of these bootstrap samples and store the estimates in a new data framed called `bstrap_fits`.

```{r}
bstrap_fits <- bstrap_samples |>
  fit()
```

c.  **Demo**: Compute 95% confidence intervals for the slope and the intercept using the `get_confidence_interval` command.

```{r}
ci_95 <- get_confidence_interval(bstrap_fits,
                        point_estimate = observed_fit,
                        level = 0.95,
                        type = "percentile")

ci_95
```

d.  **Your turn**: Verify that you get the same numbers when you manually calculate the quantiles of the slope estimates using `summarize` and `quantile`. *Pay attention to the grouping*.

```{r}
bstrap_fits |>
  filter(term == "population_share") |>
  ungroup() |>
  summarize(lower_ci = quantile(estimate, .025),
            upper_ci = quantile(estimate, .975))
```

e.  **BONUS**: You can visualize the confidence interval:

```{r}
visualize(bstrap_fits) +
  shade_confidence_interval(ci_95)
```
