---
title: "AE 13: Modeling penguins"
subtitle: "Suggested answers"
format:
  html:
    self-contained: true
---

In this application exercise we will be studying penguins.
The data can be found in the **palmerpenguins** package and we will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

Please read the following context and take a `glimpse` at the data set before we get started.

> This data set comprising various measurements of three different penguin species, namely Adelie, Gentoo, and Chinstrap.
> The rigorous study was conducted in the islands of the Palmer Archipelago, Antarctica.
> These data were collected from 2007 to 2009 by Dr. Kristen Gorman with the Palmer Station Long Term Ecological Research Program, part of the US Long Term Ecological Research Network.
> The data set is called `penguins`.

```{r}
#| label: glimpse-penguins
glimpse(penguins)
```

Our goal is to understand better how various body measurements and attributes of penguins relate to their body mass.
First, we are going to investigate the relationship between a penguins' flipper lengths and their body masses.

-   **Question:** Based on our research focus, which variable is the response variable?

`body_mass_g`

-   **Demo:** Visualize the relationship between flipper length and body mass of penguins.

```{r}
#| label: plot-bm-fl
#| warning: false

penguins |>
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()

```

# Correlation

-   **Your turn:**

    -   What is correlation? What values can correlation take?

    A measure of the strength **and** direction of the **linear** relationship between the predictor(s) and outcome.
    It can be a number between -1 and 1, inclusive.

-   **Demo:** What is the correlation between flipper length and body mass of penguins?

```{r}
#| label: correlation

penguins |>
  summarize(r = cor(flipper_length_mm, body_mass_g, use = "complete.obs"))

```

-   **Demo:** What is the coefficient of determination?

```{r}
#| label: r2-by-hand

0.8712018^2	

```

-   **Demo:** What is the slope of the regression line?

The estimated slope, as computed by hand, is \~ 49.7

```{r}
#| label: ols-by-hand

penguins |>
  summarize(
    r = cor(flipper_length_mm, body_mass_g, use = "complete.obs"),
    sy = sd(body_mass_g, na.rm = TRUE),
    sx = sd(flipper_length_mm, na.rm = TRUE),
    b1 = r * sy / sx
  )

```

# Defining, fitting, and summarizing a model

-   **Demo:** Write the **population model** below that explains the relationship between body mass and flipper length.

$$
\text{body mass (g)} = \beta_0 + \beta_1 \times \text{flipper length (mm)} + \epsilon
$$

-   **Demo:** Fit the linear regression model and display the results. Write the **fitted** model below.

```{r}
#| label: fit-bm-fl

penguins_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm, data = penguins)

tidy(penguins_fit)

```

$$
\widehat{\text{body mass(g)}} = -5780.83136 + 49.68557 \times \text{flipper length (mm)}
$$

-   **Your turn:** Interpret the slope and the intercept in the context of the data.

    -   Intercept: For a penguin with a flipper length of 0mm, we expect it to have a body mass of -5780.83g, on average.
        This is not meaningful in the context of these data because a penguin with 0mm flipper length doesn't exist.

    -   Slopes: For every 1 mm increase in flipper length, we expect a penguin's body mass to inc. by 49.69 g, on average.

-   **Your turn:** Recreate the visualization from above, this time adding a regression line to the visualization `geom_smooth(method = "lm")`.

```{r}
#| label: plot-bm-fl-line
#| message: false
#| warning: false

penguins |>
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm", color = "cornflowerblue", se = FALSE)

```

-   **Demo:** What is the coefficient of determination ($R^2$)?

75.9% of the variation in body mass (g) is explainable by the variation in flipper length (mm)

```{r}
#| label: r2-done-right

penguins_fit |>
  glance() |>
  select(r.squared)

```

-   **Demo:** What is the estimated body mass for a penguin with a flipper length of 210?
-   **Your turn:** What is the estimated body mass for a penguin with a flipper length of 100?

```{r}
#| label: predict-210-100

est_bm <- tibble(flipper_length_mm = c(210, 100))

penguins_fit |>
  predict(new_data = est_bm)

## alternative method

penguins_fit |>
  augment(new_data = est_bm)

```

# Another model

-   **Demo:** A different researcher wants to look at body weight of penguins based on the island they were recorded on. How are the variables involved in this analysis different?

Island is a categorical variable while flipper length was numeric

*Add response here*

-   **Demo:** Make an appropriate visualization to investigate this relationship below. Additionally, calculate the mean body mass by island.

```{r}
#| label: plot-bm-island
#| warning: false
#| message: false

penguins |>
  ggplot(aes(x = body_mass_g)) +
  geom_histogram() +
  facet_wrap(~ island) +
  labs(x = "Body Mass (g)",
       y = "Count",
       title = "Distribution of Penguin Body Mass by Island")

```

```{r}
#| label: mean-bm-island

penguins |>
  group_by(island) |>
  summarize(mean_bmg = mean(body_mass_g, na.rm = TRUE))

```

-   **Demo:** Change the geom of your previous plot to `geom_point()`. Use this plot to think about how R models these data.

```{r}
#| label: plot-bm-island-points
#| message: false
#| warning: false

penguins |>
  ggplot(aes(x = island, y = body_mass_g)) +
  geom_point()

```

-   **Your turn:** Fit the linear regression model and display the results. Write the **fitted** model below.

```{r}
#| label: fit-bm-island

penguins_fit_2 <- linear_reg() |>
  fit(body_mass_g ~ island, data = penguins)

tidy(penguins_fit_2)
  
```

$$
\widehat{\text{body mass (g)}} = 4716.018   -1003.115 \times \text{Dream} - 1009.645 \times \text{Torgerson}
$$

-   **Demo:** Interpret each coefficient in context of the problem.

    -   Intercept: A penguin from Biscoe island is predicted to have a body mass of 4,716.02g, on average.

    -   Slopes: A penguin from Dream island is predicted to have a body mass that is 1,003.12g lower than that of a penguin from Biscoe island, on average; a penguin from Torgerson island is predicted to have a body mass that is 1,009.65g lower than that of a penguin from Biscoe island, on average.

-   **Demo:** What is the estimated body weight of a penguin on Biscoe island?
    What are the estimated body weights of penguins on Dream and Torgersen islands?

```{r}
#| label: predict-bm-island
three_penguins <- tibble(
  island = c("Biscoe", "Dream", "Torgersen")
)
penguins_fit_2 |>
  predict(new_data = three_penguins)
```
