---
title: "AE 12: Modeling fish 🐟"
subtitle: "Suggested answers"
format:
  html:
    self-contained: true
---

For this application exercise, we will work with data on fish.
The dataset we will use, called `fish`, is on two common fish species in fish market sales.

```{r}
#| label: load-packages-data
#| message: false
library(tidyverse)
library(tidymodels)

fish <- read_csv("data/fish.csv")
```

The data dictionary is below:

| **variable**      | **description**        |
|-------------------|------------------------|
| `species`         | Species name of fish   |
| `weight`          | Weight, in grams       |
| `length_vertical` | Vertical length, in cm |
| `length_diagonal` | Diagonal length, in cm |
| `length_cross`    | Cross length, in cm    |
| `height`          | Height, in cm          |
| `width`           | Diagonal width, in cm  |

# Visualizing the model

We're going to investigate the relationship between the weights and heights of fish.

-   Create an appropriate plot to investigate this relationship. Add appropriate labels to the plot.

```{r}
#| label: plot-hw
fish |>
  ggplot(aes(x = height, y = weight)) +
  geom_point() +
  labs(x = "Height (cm)",
       y = "Weight (g)",
       title = "Fish Weight vs. Height")
```

-   If you were to draw a a straight line to best represent the relationship between the heights and weights of fish, where would it go? Why?

Positive, y-intercept below 0

-   Now, let R draw the line for you.

```{r}
#| label: plot-hw-lm
#| message: false
fish |>
  ggplot(aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") +
  labs(x = "Height (cm)",
       y = "Weight (g)",
       title = "Fish Weight vs. Height")
```

-   What types of questions can this plot help answer?

We can now describe the relationship between fish height and weight, and we could also make predictions of a fish's weight given its height.

-   We can use this line to make predictions. Predict what you think the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm. Which prediction is considered extrapolation?

10cm: Between 310-325g

15cm: 625g

20cm: 920g - this prediction is considered an extrapolation because we don't observe any fish with weights \> 19.5cm in our dataset

-   What is a residual?

Observed - predicted; in statistical notation, this is $y - \widehat{y}$

# Model fitting

-   **Demo:** Fit a model to predict fish weights from their heights.

```{r}
#| label: fit-model
fish_fit <- linear_reg() |>
  fit(weight ~ height, data = fish)

tidy(fish_fit)
```

-   Predict what the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm using this model.

```{r}
#| label: predict-by-hand
y_10 <- -288.415 + (60.916 * 10)
y_15 <- -288.415 + (60.916 * 15)
y_20 <- -288.415 + (60.916 * 20)

predict <- c(y_10, y_15, y_20)
predict
```

-   **Demo:** Calculate predicted weights for all fish in the data and visualize the residuals under this model.

```{r}
#| label: residuals
#| message: false
#| warning: false
fish_augment <- augment(fish_fit, new_data = fish)

ggplot(fish_augment, aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") +
  geom_segment(aes(xend = height, yend = .pred), color = "grey") +
  geom_point(aes(y = .pred), shape = "circle open") +
  theme_minimal() +
  labs(x = "Height (cm)",
       y = "Weight (g)",
       title = "Fish Weight vs. Height",
       subtitle = "Residuals overlaid")
```

# Model summary

-   **Demo:** Display the model summary including estimates for the slope and intercept along with measurements of uncertainty around them. Show how you can extract these values from the model output.

```{r}
#| label: model-summary
tidy_fit <- tidy(fish_fit)
int = tidy_fit$estimate[1]
int
slope = tidy_fit$estimate[2]
slope
```

-   **Demo:** Write out your model using mathematical notation.

$\widehat{\text{weight}} = -288.415 + 60.916 \times \text{height}$ - Fitted model

$\text{weight} = \beta_0 + \beta_1 \times \text{height} + \epsilon$ - True population model

# Correlation

We can also assess correlation between two quantitative variables.

-   What is correlation? What are values correlation can take?

Correlation is a measure of the strength and direction of the *linear* relationship between two variables; it is a number between -1 and 1, inclusive.

-   **Demo:** What is the correlation between heights and weights of fish?

```{r}
#| label: correlation
fish |>
  summarize(r = cor(height, weight))
```

# Adding a third variable

-   **Demo:** Does the relationship between heights and weights of fish change if we take into consideration species? Plot two separate straight lines for the Bream and Roach species.

```{r}
#| label: plot-hws-lm
#| warning: false
#| message: false
fish |>
  ggplot(aes(x = height, y = weight, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Height (cm)",
       y = "Weight (g)",
       title = "Fish Weight vs. Height")
```

# Fitting other models

-   **Demo:** We can fit more models than just a straight line. Change the following code below to read `method = "loess"`. What is different from the plot created before?

```{r}
#| label: plot-hw-loess
#| warning: false
#| message: false
ggplot(fish, 
       aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(
    title = "Weights vs. heights of fish",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```
