---
title: "AE 12: Modelling penguins"
subtitle: "Suggested answers"
---

In this application exercise we will be studying penguins.
The data can be found in the **palmerpenguins** package and we will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

You've seen the `penguins` data far too much at this point (sorry!!!), but we're going to bring it back one more time.

```{r}
#| label: glimpse-penguins
glimpse(penguins)
```

# Part 1

Our goal is to understand better **how various body measurements and attributes of penguins relate to their body mass**.
First, we are going to investigate the relationship between a penguins' flipper lengths and their body masses.

-   *Based on our research focus, body mass is the response variable.*

**Task 1 - Exploratory Data Analysis:** Visualize the relationship between flipper length and body mass of penguins.

```{r}
#| label: plot-bm-fl

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

## Correlation

**Task 2 - Complete the following**

-   **Question:** What is correlation?
    What values can correlation take?

    *Strength and direction of a linear relationship. It's bounded by -1 and 1.*

    ::: callout-note
    *Are you good at guessing correlation?* [Give it a try](https://www.guessthecorrelation.com) with this game!
    :::

-   **Code:** What is the correlation between flipper length and body mass of penguins?

```{r}
#| label: correlation

# option 1
penguins |>
  summarize(r = cor(flipper_length_mm, body_mass_g, use = "complete.obs"))

# option 2
penguins |>
  drop_na(flipper_length_mm, body_mass_g) |>
  summarize(r = cor(flipper_length_mm, body_mass_g))
```

## Defining, fitting, and summarizing a model

**Task 3 (Demo) :** Write the [*population model*]{.underline} (model for the *true values)* below that explains the relationship between body mass and flipper length.

$$
{body mass} = \beta_0 + \beta_1 \times flipper length + \epsilon
$$**Task 4:** Fit the linear regression model and display the results.
[*Write the estimated model output below!*]{.underline}

```{r}
#| label: fit-bm-fl

bm_fl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm, data = penguins)

tidy(bm_fl_fit)
```

$$
\widehat{body mass} = -5780 + 49.7 \times flipper length 
$$

**Task 5:** Interpret the slope and the intercept in the context of the data.

-   *Intercept*: Penguins with 0 flipper length are expected, on average, to weigh -5,781 grams.

-   *Slope*: For each additional millimeter of a penguin;s flipper length, the weight of their penguin is expected to be higher, on average, by 49.7 grams.

**Task 6:** Recreate the visualization from above, this time adding a regression line to the visualization `geom_smooth(method = "lm")`.

```{r}
#| label: plot-bm-fl-line

ggplot(
  penguins,
  aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_point() +
  geom_smooth(method = "lm")
```

**Task 7 (Demo):** What is the estimated body mass for a penguin with a flipper length of 210?

```{r}
#| label: predict-210

penguin_flipper_210 <- tibble(flipper_length_mm = 210)

bm_fl_fit |>
  predict(new_data = penguin_flipper_210)
```

**Task 8:** What is the estimated body mass for a penguin with a flipper length of 100?
Add code to find it!
Is there anything weird about making this prediction?

```{r}
#| label: predict-100

penguin_flipper_100 <- tibble(flipper_length_mm = 100)

bm_fl_fit |>
  predict(new_data = penguin_flipper_100)
```

We are extrapolating when we take flipper length = 100: it is outside of the data range.

# Part 2: Another model

**Task 9:** A different researcher wants to look at body weight of penguins based on the island they were recorded on.

-   **Question:** How are the variables involved in this analysis different?

    The predictor variable is categorical, not quantitative!

-   **Code:** Make an appropriate visualization to investigate this relationship below.
    Additionally, calculate the mean body mass by island.

```{r}
#| label: plot-bm-island


penguins |>
  ggplot(aes(x = island, y = body_mass_g)) +
  geom_boxplot()
```

```{r}
#| label: mean-bm-island

penguins |>
  group_by(island) |>
  summarize(mean_bm = mean(body_mass_g, na.rm = TRUE))
```

**Task 10:** Change the geom of your previous plot to `geom_point()`.
Use this plot to think about how R models these data.

```{r}
#| label: plot-bm-island-points

penguins |>
  ggplot(aes(x = island, y = body_mass_g)) +
  geom_point()
```

**Task 11:** Fit the linear regression model and display the results.
Write the estimated model output below.

```{r}
#| label: fit-bm-island


bm_island_fit <- linear_reg() |>
  fit(body_mass_g ~ island, data = penguins)

tidy(bm_island_fit)
```

$$
\widehat{bodymass} = 4716 -1003 \times dream -1010 \times torgersen
$$

## Interpreting Categorical Predictors

**Task 12:** Fill in the blanks.

-   The baseline island is Biscoe.
-   Intercept: Penguins from Biscoe island are expected to weigh, on average, 4716 grams.
-   Slopes:
    -   Penguins from Dream are expected to weigh, on average, 1003 grams less than those from Biscoe.

    -   Penguins from Torgersen island are expected to weigh, on average, 1010 grams less than those from Biscoe.

**Task 13:** What is the estimated body weight of a penguin on Biscoe island?
What are the estimated body weights of penguins on Dream and Torgersen islands?

```{r}
#| label: predict-bm-island


three_penguins <- tibble(
  island = c("Biscoe", "Dream", "Torgersen")
)

bm_island_fit |>
  predict(new_data = three_penguins)
```
