---
title: "AE 14: Modeling Housing Prices 🏠"
subtitle: "Suggested answers"
format:
  html:
    self-contained: true
---

In this application exercise we will be studying housing prices.
The dataset is a cleaned version of publicly available real estate data.
We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(modeldata)

ames <- modeldata::ames
```

We will use the `ames_housing` dataset from the **modeldata** package.

Before we use the dataset, we'll make a few transformations to it.

-   **Your turn:** Review the code below with your neighbors and write a summary of the data transformation pipeline.

-   Selected for a subset of relevant columns

-   Created a new variable, `home_type`, by consolidating the levels of the og.
    factor variable, `Bldg_Type`; then, got rid of `Bldg_Type`

-   Renamed some columns

-   Subset df to homes of type House, Townhouse, or Duplex

```{r}
#| label: data-prep
data(ames)
housing <- ames |>
  select(Sale_Price, Gr_Liv_Area, Bldg_Type, Bedroom_AbvGr, Paved_Drive, Exter_Cond) |>
  mutate(home_type = fct_collapse(Bldg_Type,
    "House" = c("OneFam", "TwnhsE"),
    "Townhouse" = "Twnhs",
    "Duplex" = "Duplex"
  )) |>
  select(-Bldg_Type) |>
  rename(price = Sale_Price, sqft = Gr_Liv_Area, bedrooms = Bedroom_AbvGr) |>
  filter(home_type %in% c("House", "Townhouse", "Duplex"))
```

Here is a glimpse at the data:

```{r}
#| label: glimpse-housing
glimpse(housing)
```

# Get to know the data

-   **Your turn:** What is a typical house price in this dataset? What are some common square footage values? What types of homes are most common? Additionally, explore at least 1-2 other features that could be interesting. Share your findings!

```{r}
#| label: explore-housing
#| layout-ncol: 2
#| message: false

ggplot(housing, aes(x = price)) +
  geom_boxplot()

ggplot(housing, aes(x = sqft)) +
  geom_histogram()

housing |>
  count(home_type)
```

# Price vs. square footage

How can we use square footage to model/predict pricing?
Here is the model:

```{r}
#| label: price-sqft-fit
price_sqft_fit <- linear_reg() |>
  fit(price ~ sqft, data = housing)

tidy(price_sqft_fit)
```

And here is the model visualized:

```{r}
#| label: price-sqft-plot
#| message: false
#| warning: false
ggplot(housing, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")
```

-   **Your turn:** Write the fitted equation of the model in mathematical notation. Then, interpret the intercept and slope.

$$
\widehat{price} = 11430.1960    +113.7605   \times sqft
$$ For a home with 0 sq.
ft., we predict the price to be \$11430.20, on average.
For each additional sq.
ft, we predict the price of a home to increase by \$113.76, on average.

# Price vs. home type

```{r}
#| label: price-type-fit
price_type_fit <- linear_reg() |>
  fit(price ~ home_type, data = housing)

tidy(price_type_fit)
```

-   **Your turn:** Write the fitted equation of the model in mathematical notation. Then, interpret the intercept and each coefficient in context.

$$
\widehat{price} = 185469.48 - 45660.54 \times Duplex - 49535.42 \times Townhouse
$$ We predict the price of a house to be \$185,469.48, on average.
We predict that the price of a duplex will be \$45,660.54 lower than the price of a house, on average.
We predict that the price of a townhouse will be \$49,535.42 lower than the price of a house, on average.

# Price vs. square footage and home type

Now, let's fit a model that use both variables!

## Main effects model

The *main effects model* is another name for the additive model.
Here is the model:

```{r}
#| label: price-main-fit
price_main_fit <- linear_reg() |>
  fit(price ~ sqft + home_type, data = housing)

tidy(price_main_fit)
```

-   **Your turn:** Write the fitted equation of the model in mathematical notation. Then, interpret the intercept and each slope coefficient.

$$
\widehat{price} = 13394.8010 + 114.5305 \times sqft - 63251.4927 \times Duplex - 20306.3316 \times Townhouse
$$ We predict the price of a house with 0 sq.
ft. to be \$13,394.80, on average.
All else equal, for each additional sq.
ft., we predict the price of a home to increase by \$114.53, on average.
All else equal, we predict the price of a Duplex to be \$63,251.50 lower than the price of a house, on average.

## Interaction effects model

Now, we will fit an interaction effects model.

**Task:** Write code to fit an interaction effects model predicting price from square footage and home type.

```{r}
#| label: price-int-fit
model_int <- linear_reg() |>
  fit(price ~ sqft * home_type, data = housing)

tidy(model_int)
```

**Task:** Write the fitted equation using mathematical notation.

$$
\begin{aligned}
\widehat{price} = 9087.38517 + 117.39744 \times sqft + \\
57547.32571 \times Duplex + 13970.23460 \times Townhouse \\
- 73.21084 \times (sqft *Duplex) -26.89557 (sqft * Townhouse)
\end{aligned}
$$

**Task:** Write the fitted equations for each home type.

Duplex

$$
\widehat{price} = (9087.38517 + 57547.32571) + (117.39744 - 73.21084) \times sqft
$$ House

$$
\widehat{price} = 9087.38517 + 117.39744 \times sqft
$$ Townhouse $$
\widehat{price} = (9087.38517 + 13970.23460) + (117.39744 -26.89557) \times sqft
$$

# Model Comparison

So, we fit multiple models - how do we know which one is better?

Using `glance()`, report each of the above models' *adjusted* $R^2$ values.

Which model is the best fit?
Which is the worst?

```{r}
#| label: model-glance

glance(model_int)$adj.r.squared
glance(price_main_fit)$adj.r.squared
glance(price_sqft_fit)$adj.r.squared
glance(price_type_fit)$adj.r.squared

```

# One more model?

**Task:** Try adding one more variable present in the data frame to the "best" model chosen above.
Does it make a difference in *adjusted* $R^2$?
What about plain old $R^2$?

```{r}
#| label: new-model

glimpse(housing)

housing_new_fit <- linear_reg() |>
  fit(price ~ sqft * home_type + bedrooms, data = housing)

tidy(housing_new_fit)

glance(housing_new_fit)$adj.r.squared
glance(housing_new_fit)$r.squared
glance(model_int)$r.squared
glance(model_int)$r.squared

```
