---
title: "AE 18: Stocks!!"
---

Today we'll explore the question "How do stock prices of Apple and Microsoft relate to each other?"

# Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
```

# Data

The data for this application exercise was originally gathered using the [**tidyquant**](https://business-science.github.io/tidyquant/) R package.
It features Apple and Microsoft stock prices from January 1st 2020 to December 31st 2021.

First, let's load the data:

```{r}
#| label: load-data
#| message: false

stocks <- read_csv("data/stocks.csv")
```

To keep things simple, we'll work with a subset of the data, stock prices in January 2020.

```{r}
stocks_jan2020 <- stocks |>
  filter(month(date) == 1 & year(date) == 2020)

glimpse(stocks_jan2020)
```

These are daily data.
For each stock, we have several variables, including:

-   `Open`: price at the start of the trading day;
-   `High`: highest price during the trading day;
-   `Low`: lowest price during the trading day;
-   `Close`: price at the end of the trading day;

We will focus on using the Microsoft **open** to predict the Apple **open**.

# Visualize

**Your turn**: Plot the data and the line of best fit.

```{r}
#| message: false
ggplot(stocks_jan2020, aes(x = MSFT.Open, y = AAPL.Open)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

# Point Estimation

a.  **Your turn**: Fit the linear model to these data:

```{r}
observed_fit <- stocks_jan2020 |>
  specify(AAPL.Open ~ MSFT.Open) |>
  fit()
observed_fit
```

::: callout-note
This gives the exact same numbers that you get if you use `linear_reg() |> fit()`, but we need this new syntax because it plays nice with the tools we have for confidence intervals and hypothesis tests.
I know, I hate it too, but it's the way it is.
:::

b.  **Your turn**: Typeset the equation for the model fit:

$$
\widehat{AAPL} = 3.31 + 0.45\times MSFT
$$

c.  **Your turn**: Interpret the slope and the intercept estimates:

    -   If Microsoft opened at \$0, we predict that Apple opens at \$3.31 on average;
    -   A \$1 increase in Microsoft predicts ia associated with a \$0.45 increase in Apple, on average.

# Interval Estimation

a.  **Demo**: Using seed `8675309`, generate `500` bootstrap samples, and store them in a new data frame called `bstrap_samples`.

```{r}
set.seed(8675309)
bstrap_samples <- stocks_jan2020 |>
  specify(AAPL.Open ~ MSFT.Open) |>
  generate(reps = 500, type = "bootstrap")
```

b.  **Demo**: Fit a linear model to each of these bootstrap samples and store the estimates in a new data framed called `bstrap_fits`.

```{r}
bstrap_fits <- bstrap_samples |> 
  fit()
```

c.  **Your turn**: Use `linear_reg() |> fit(...)` to fit a linear model to bootstrap sample number *347*, and verify that you get the same estimates as the ones contained in `bstrap_fits`.

```{r}
replicate_347 <- bstrap_samples |>
  filter(
    replicate == 347
  )

linear_reg() |>
  fit(AAPL.Open ~ MSFT.Open, data = replicate_347) |>
  tidy()

bstrap_fits |>
  filter(replicate == 347)
```

::: callout-note
The only point I'm making here is that this new bootstrap code is not performing a fundamentally new task.
It's performing an old task (fitting the linear model), but it's repeating it A LOT. So the numbers you get here are not mysterious.
They're numbers you already know how to compute.
:::

d.  **Demo**: Compute 90% confidence intervals for the slope and the intercept using the `get_confidence_interval` command.

```{r}
ci_90 <- get_confidence_interval(
  bstrap_fits,
  point_estimate = observed_fit,
  level = 0.90,
  type = "percentile"
)
ci_90
```

e.  **Your turn**: Verify that you get the same numbers when you manually calculate the quantiles of the slope estimates using `summarize` and `quantile`. *Pay attention to the grouping*.

```{r}
bstrap_fits |>
  ungroup() |>
  group_by(term) |>
  summarize(
    lower_ci = quantile(estimate, 0.05),
    upper_ci = quantile(estimate, 0.95)
  )
```

::: callout-note
Same point as before.
There's no magic here.
`get_confidence_interval` is just a convenient way of doing something that you already knew how to do.s
:::

f.  **BONUS**: You can visualize the confidence interval:

```{r}
visualize(bstrap_fits) + 
  shade_confidence_interval(ci_90)
```

# Hypothesis Testing

Let's consider the hypotheses:

$$
H_0:\beta_1=0\quad vs\quad H_A: \beta_1\neq 0.
$$ The null hypothesis corresponds to the claim that Microsoft and Apple stock prices are uncorrelated.
Probably bogus.

a.  Simulate and plot the null distribution for the slope:

```{r}
#| message: false
set.seed(20241118)
null_dist <- stocks_jan2020 |>
  specify(AAPL.Open ~ MSFT.Open) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  fit()

null_dist |>
  filter(term == "MSFT.Open") |>
  ggplot(aes(x = estimate)) + 
  geom_histogram()
```

b.  Add a vertical line to your plot indicating the point estimate of the slope from your original data analysis:

```{r}
visualize(null_dist) +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```

c.  Compute the $p$-value for this test and interpret it:

```{r}
#| warning: false
null_dist |>
  get_p_value(obs_stat = observed_fit, direction = "two-sided")
```

**Answer**: If the null were true and $\beta_1=0$ (meaning Apple and Microsoft stock prices are uncorrelated), the probability of getting an estimate as or more extreme than the one we got ($b_1=0.45$) is essentially 0.
So to hell with the null!
We reject it.
Evidence is sufficient to conclude that $\beta_1$, whatever it is, is not equal to zero.

# 
