---
title: "AE 16: Houses in Duke Forest"
---

In this application exercise, we use bootstrapping to construct confidence intervals.

## Packages

We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively, and the **openintro** package for the data, and the **knitr** package for formatting tables.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(openintro)
library(knitr)
```

## Data

The data are on houses that were sold in the Duke Forest neighborhood of Durham, NC around November 2020.
It was originally scraped from Zillow, and can be found in the `duke_forest` data set in the **openintro** R package.

```{r}
#| label: glimpse-data
glimpse(duke_forest)
```

## Bootstrap confidence interval

### 1. Calculate the observed fit (slope)

```{r}
#| label: observed-fit
#| eval: false
observed_fit <- duke_forest |>
  specify(____ ~ _____) |>
  fit()

observed_fit
```

### 2. Take *n* bootstrap samples and fit models to each one.

Fill in the code, then set `eval: true` .

```{r}
#| label: bootstrap
#| eval: false
n = 100
set.seed(1234)

boot_fits <- duke_forest |>
  _____(price ~ area) |>
  ______(reps = _____, type = ______) |>
  fit()

boot_fits
```

-   Why do we set a seed before taking the bootstrap samples?

-   Make a histogram of the bootstrap samples to visualize the bootstrap distribution.

```{r}
#| label: boot-hist
#| eval: false


boot_fits |>
  filter(_____  == _______) |>
  ggplot(aes(x = ______ )) +
  geom_histogram()
```

### 3. Compute the 95% confidence interval as the middle 95% of the bootstrap distribution

```{r}
#| label: calc-ci
#| eval: false

get_confidence_interval(
  boot_fits, 
  point_estimate = _____, 
  level = ______,
  type = ______
)
```

## Changing confidence level

### Modify the code from Step 3 to create a 90% confidence interval.

```{r}
#| label: 90-ci
#| eval: false
get_confidence_interval(
  boot_fits, 
  point_estimate = observed_fit, 
  level = ____,
  type = "percentile"
)
```

### Modify the code from Step 3 to create a 99% confidence interval.

```{r}
#| label: 99-ci
#| eval: false
get_confidence_interval(
  boot_fits, 
  point_estimate = observed_fit, 
  level = _____,
  type = "percentile"
)
```

-   Which confidence level produces the most accurate confidence interval (90%, 95%, 99%)?
    Explain

-   Which confidence level produces the most precise confidence interval (90%, 95%, 99%)?
    Explain

-   If we want to be very certain that we capture the population parameter, should we use a wider or a narrower interval?
    What drawbacks are associated with using a wider interval?
