---
title: "AE 14: Modeling Housing Prices 🏠"
---

In this application exercise we will be studying housing prices.
The dataset is a cleaned version of publicly available real estate data.
We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(modeldata)
```

We will use the `ames_housing` dataset from the **modeldata** package.

Before we use the dataset, we'll make a few transformations to it.

-   **Your turn:** Review the code below with your neighbors and write a summary of the data transformation pipeline.

```{r}
#| label: data-prep
data(ames)
housing <- ames |>
  select(Sale_Price, Gr_Liv_Area, Bldg_Type, Bedroom_AbvGr, Paved_Drive, Exter_Cond) |>
  mutate(home_type = fct_collapse(Bldg_Type,
    "House" = c("OneFam", "TwnhsE"),
    "Townhouse" = "Twnhs",
    "Duplex" = "Duplex"
  )) |>
  select(-Bldg_Type) |>
  rename(price = Sale_Price, sqft = Gr_Liv_Area, bedrooms = Bedroom_AbvGr) |>
  filter(home_type %in% c("House", "Townhouse", "Duplex"))
```

Here is a glimpse at the data:

```{r}
#| label: glimpse-housing
glimpse(housing)
```

# Get to know the data

-   **Your turn:** What is a typical house price in this dataset? What are some common square footage values? What types of homes are most common? Additionally, explore at least 1-2 other features that could be interesting. Share your findings!

```{r}
#| label: explore-housing
#| layout-ncol: 2

# add code here
```

# Price vs. square footage

How can we use square footage to model/predict pricing?
Here is the model:

```{r}
#| label: price-sqft-fit
price_sqft_fit <- linear_reg() |>
  fit(price ~ sqft, data = housing)

tidy(price_sqft_fit)
```

And here is the model visualized:

```{r}
#| label: price-sqft-plot
ggplot(housing, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")
```

-   **Your turn:** Write the fitted equation of the model in mathematical notation. Then, interpret the intercept and slope.

$$
add~equation~here
$$

# Price vs. home type

```{r}
#| label: price-type-fit
price_type_fit <- linear_reg() |>
  fit(price ~ home_type, data = housing)

tidy(price_type_fit)
```

-   **Your turn:** Write the fitted equation of the model in mathematical notation. Then, interpret the intercept and each coefficient in context.

$$
add~equation~here
$$

# Price vs. square footage and home type

Now, let's fit a model that use both variables!

## Main effects model

The *main effects model* is another name for the additive model.
Here is the model:

```{r}
#| label: price-main-fit
price_main_fit <- linear_reg() |>
  fit(price ~ sqft + home_type, data = housing)

tidy(price_main_fit)
```

-   **Your turn:** Write the fitted equation of the model in mathematical notation. Then, interpret the intercept and each slope coefficient.

$$
add~equation~here
$$

## Interaction effects model

Now, we will fit an interaction effects model.

**Task:** Write code to fit an interaction effects model predicting price from square footage and home type.

```{r}
#| label: price-int-fit

#add code here
```

**Task:** Write the fitted equation using mathematical notation.

$$
add~equation~here
$$

**Task:** Write the fitted equations for each home type.

$$
add~equation~here
$$
$$
add~equation~here
$$
$$
add~equation~here
$$

# Model Comparison

So, we fit multiple models - how do we know which one is better?

Using `glance()`, report each of the above models' *adjusted* $R^2$ values.

Which model is the best fit?
Which is the worst?

```{r}
#| label: model-glance

# add code here
```

# One more model?

**Task:** Try adding one more variable present in the data frame to the "best" model chosen above.
Does it make a difference in *adjusted* $R^2$? What about plain old $R^2$?

```{r}
#| label: new-model

# add code here
```
