---
title: "AE 13: Modeling housing prices"
subtitle: "Suggested answers"
---

::: callout-important
These are suggested answers.
This document should be used as reference only, it's not designed to be an exhaustive key.
:::

In this application exercise we will be studying housing prices.
The dataset is a cleaned version of publicly available real estate data.
We will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(modeldata)
```

We will use the `ames_housing` dataset from the **modeldata** package.

Before we use the dataset, we'll make a few transformations to it.

-   **Your turn:** Review the code below with your neighbor and write a summary of the data transformation pipeline.

*Add response here.*

```{r}
#| label: data-prep
data(ames)
housing <- ames |>
  select(Sale_Price, Gr_Liv_Area, Bldg_Type, Bedroom_AbvGr, Paved_Drive, Exter_Cond) |>
  mutate(home_type = fct_collapse(Bldg_Type,
    "House" = c("OneFam", "TwnhsE"),
    "Townhouse" = "Twnhs",
    "Duplex" = "Duplex"
  )) |>
  select(-Bldg_Type) |>
  rename(price = Sale_Price, sqft = Gr_Liv_Area, bedrooms = Bedroom_AbvGr) |>
  filter(home_type %in% c("House", "Townhouse", "Duplex"))
```

Here is a glimpse at the data:

```{r}
#| label: glimpse-housing
glimpse(housing)
```

# Get to know the data

-   **Your turn:** What is a typical house price in this dataset? What are some common square footage sizes? What types of homes are most common? Additionally, explore at least 1-2 other features that could be interesting. Share your findings!

```{r}
#| label: explore-housing
#| layout-ncol: 2
ggplot(housing, aes(x = price)) +
  geom_histogram(binwidth = 20000)

ggplot(housing, aes(x = sqft)) +
  geom_histogram(binwidth = 250)

ggplot(housing, aes(x = home_type)) +
  geom_bar()
```

# Price vs. square footage

How can we use square footage to model/predict pricing?
Here is the model:

```{r}
#| label: price-sqft-fit
price_sqft_fit <- linear_reg() |>
  fit(price ~ sqft, data = housing)

tidy(price_sqft_fit)
```

And here is the model visualized:

```{r}
#| label: price-sqft-plot
ggplot(housing, aes(x = sqft, y = price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")
```

-   **Your turn:** Write the equation of the model in mathematical notation. Then, interpret the intercept and slope.

$$ \widehat{price} = 11430.20 + 113.76 \times sqft $$

-   **Intercept**: On average, we expect a home that is 0 square feet to cost \$11430.
    This does not make sense in the context of the data - we won't have a home that is 0 sq.
    ft.

-   **Slope:** On average, for every 1 additional square foot, we expect hosue price to increase by \$113.76

# Price vs. home type

```{r}
#| label: price-type-fit
price_type_fit <- linear_reg() |>
  fit(price ~ home_type, data = housing)

tidy(price_type_fit)
```

-   **Your turn:** Write the equation of the model in mathematical notation. Then, interpret the intercept and each coefficient in context.

$$
\widehat{price} = 185469.48 - 45660.54 \times Duplex - 49535.42 \times Townhouse
$$

-   **Intercept:** On average, we expect houses to cost 185469.48.

-   **Slope for Duplex:** On average, we expect duplexes to cost 45660 dollars less than a house.

# Price vs. square footage and home type

Now, let's make some model that use both variables!

## Main effects model

The *main effects model* is another name for the additive model.
We fit the models below and wrote the model in math notation.

```{r}
#| label: price-main-fit
price_main_fit <- linear_reg() |>
  fit(price ~ sqft + home_type, data = housing)

tidy(price_main_fit)
```

$$
\widehat{price} = 13394.8 + 114.5 \times sqft - 20306 \times Townhouse - 63251 \times Duplex
$$

**Task:** Write the model equations for each home type.
Provide interpretations of the coefficients.

**House:** \$ \widehat{price} = 13394.8 + 114.5 \times sqft \$

-   Intercept: On average, we expect a house with 0 square feet to cost \$13394.8.
-   Slope: On average, for houses, we expect every one additional square foot to correspond to an additional price of \$114.5.

**Townhouse:** \$ \widehat{price} = (13394.8 - 20306) + 114.5 \times sqft \$

-   Intercept: On average, we expect a townhouse with 0 square feet to cost \$-6911.2. This does not make sense in the context of the data - townhouses cannot be 0 sqft and prices cannot be negative.
-   Slope: On average, for townhouses, we expect every one additional square foot to correspond to an additional price of \$114.5.

**Duplex:** \$ \widehat{price} = (13394.8 - 63251) + 114.5 \times sqft \$

-   Intercept: On average, we expect a house with 0 square feet to cost -\$49856.2. 49856.2. This does not make sense in the context of the data - duplexes cannot be 0 sqft and prices cannot be negative.
-   Slope: On average, for duplexes, we expect every one additional square foot to correspond to an additional price of \$114.5.

**NOTE**: Slopes are the SAME for all types!!

## Interaction effects model

Now, we will fit an interaction effects model.

**Task:** Write code to fit an interaction effects model predicting price from square feet and home type.

```{r}
#| label: price-int-fit

price_inter_fit <- linear_reg() |>
  fit(price ~ sqft  * home_type, data = housing)

tidy(price_inter_fit)
```

**Task:** Write the model output using mathematical notation.

$$
\widehat{price} = 9087 + 117 \times sqft + 139970 \times Townhouse + 57547 \times Duplex - 73 \times sqft \times Duplex - 27 \times sqft \times Townhouse
$$

**Task:** Write the model equations for each home type.

**House:** \$ \widehat{price} = 9087 + 117 \times sqft \$

**Townhouse:** \$ \widehat{price} = (9087 + 139970) + (117 - 27) \times sqft \$

**Duplex:** \$ \widehat{price} = (9087 + 57547) + (117 - 73) \times sqft \$

# Model Comparison

So, we fit multiple models - how do we know which one is better?

We will dive into this tomorrow, but there is a value called *adjusted* $R^2$ that lets us compare models.
Higher values are better, lower are worse.
You can `glance()` at a model fit to see the *adjusted* $R^2$ values.

Which model is the best fit?
Which is the worst?

```{r}
#| label: model-glance
glance(price_main_fit)
glance(price_inter_fit)
```

Adjusted R squared is a higher for the interaction effects model - this is the better model!
# One more model?

**Task:** Try adding one more variable to your chosen model from above.
Does it make a difference in *adjusted* $R^2$?

```{r}
#| label: new-model

price_inter_bed_fit <- linear_reg() |>
  fit(price ~ sqft  * home_type + bedrooms, data = housing)

glance(price_inter_bed_fit)
```

Adding the number of bedrooms greatly boosts the adjusted r squared value - this is a preferrable model!
