---
title: "AE 20: Final 'Live Coding' Exam Practice 🎬💳"
---

## Packages

Use **only** the following packages for this exam:

```{r}
#| message: false
library(tidyverse)
library(tidymodels)
```

# Part 1 - Movies

We will work with data from the Internet Movie Database (IMDB). Specifically, the data were a random sample of movies released between 1980 and 2020.

The variables and their descriptions in the raw movies dataset are as
follows:

| Variable          | Description                                                             |
|-------------------|-------------------------------------------------------------------------|
| `name`            | name of the movie                                                       |
| `score`           | IMDB user rating                                                        |
| `runtime`         | duration of the movie                                                   |
| `genre`           | main genre of the movie.                                                |
| `rating`          | rating of the movie (R, PG, etc.)                                       |
| `release_country` | release country                                                         |
| `release_date`    | release date (YYYY-MM-DD)                                               |
| `budget`          | the budget of a movie (some movies don’t have this, so it appears as 0) |
| `gross`           | revenue of the movie                                                    |
| `votes`           | number of user votes                                                    |
| `year`            | year of release                                                         |
| `director`        | the director                                                            |
| `writer`          | writer of the movie                                                     |
| `star`            | main actor/actress                                                      |
| `country`         | country of origin                                                       |
| `company`         | the production company                                                  |

## Question 1: Clean and recode

The data you’ll use for this question is in the `data` folder of your
repository, and it’s called `movies-raw.csv`. The goal of the question
is to “clean” this raw data and save a new version of it as
`movies-processed.csv`. The following parts walk you through what you
need to do to clean the data.

**a.** Read the dataset called `movies-raw.csv` and save it as
`movies_raw`.

```{r}
#| label: q1_part_a
movies_raw <- read_csv("data/movies-raw.csv")
```

**b.** In a single pipeline,

- remove the character string `" mins"` from `runtime` and convert
  `runtime` to numeric,
- recode the levels of `release_country` to `"United States"` and
  `"Not United States"`, in that order,
- recode the levels of `genre` to `"Action"`, `"Comedy"`, `"Drama"`,
  `"Horror"`, and `"Other"`, in that order, and
- save the resulting data frame as `movies_processed`.

```{r}
#| label: q1_part_b
movies_processed <- movies_raw |>
  mutate(
    runtime = as.numeric(str_remove(runtime, " mins")),
    release_country = if_else(release_country == "United States",
                              "United States",
                              "Not United States"),
    release_country = fct_relevel(release_country, "United States", "Not United States"),
    genre = if_else(!genre %in% c("Action", "Comedy", "Drama", "Horror"),
                     "Other",
                     genre),
    genre = fct_relevel(genre, "Action", "Comedy", "Drama", "Horror", "Other")
  ) 
```

**c.** In a single pipeline, calculate the cutoff for the top 20th
percentile of movie `score`s and store it as `top_20_cutoff` as a single
numeric value.

> [!TIP]
>
> You can check if you’ve done this correctly by typing `top_20_cutoff`
> in the Console and checking the output, it should look like the
> following:
>
>     > top_20_cutoff
>     80% 
>     7.2

```{r}
#| label: q1_part_c
top_20_cutoff <- movies_processed |>
  summarize(top_20_cutoff = quantile(score, 0.8)) |>
  pull(top_20_cutoff)
```

**d.** In a single pipeline, update `movies_processed` to include a column called `percentile` with 2 possible values: "Top 20th", if the movie's `score` is $\geq$ `top_20_cutoff`, and "Bottom 80th", if the movie's `score` is < `top_20_cutoff`. You should ensure that `percentile` is a factor variable coded such that "Bottom 80th" is the baseline level.

```{r}
#| label: q1_part_d
movies_processed <- movies_processed |>
  mutate(
    percentile = if_else(score >= top_20_cutoff, "Top 20th", "Bottom 80th"),
    percentile = fct_relevel(percentile, "Bottom 80th", "Top 20th")
  )
```

Regardless of your solution to the above question, run the chunk below to load a pre-saved `movies_processed` for use in subsequent questions. You will need to change the YAML setting to `#| eval: true` prior to rendering your `.qmd` file.

```{r}
#| label: intermed-data-1
#| eval: false
movies_processed <- read_csv("data/movies-processed.csv")
```

## Question 2: Model

**a.** Suppose that a movie studio makes decisions on whether to produce
a movie or not based on whether they think it will score in the top 20th
percentile of IMDB scores. Help them build a model to aid in their
decision making.

- Split the data into training (75%) and testing (25%) subsets 

- Fit a model predicting whether the movie is in the Top 20th
  percentile of scores based on its runtime, genre, and the interaction
  of these two predictors.

- Display a tidy output of the model.

- Interpret the intercept in the context of these data with respect to $\widehat{p}$.

For an action movie with a runtime of 0 minutes, we predict the probability of that movie scoring in the top 20th percentile to be $\frac{e^{-11.41}}{1 + e^{-11.41}} \approx 1.11\times10^{-05}$, on average.

```{r}
#| label: q2_part_a
set.seed(847)
movies_split <- initial_split(movies_processed)
movies_train <- training(movies_split)
movies_test <- testing(movies_split)

percentile_fit <- logistic_reg() |>
  fit(as.factor(percentile) ~ runtime * genre, data = movies_train)

tidy(percentile_fit)

exp(-11.410900184) / (1 + exp(-11.410900184))
```

**b.** Based on your model from part (a), calculate the predicted values
of `percentile` for movies in your testing dataset. Then, in a single pipeline,
calculate the false positive and false negative rates for this model.
Explicitly state in your narrative the false positive and false negative
rates.

The false positive rate is ~4.3%, and the false negative rate is ~75%.

```{r}
#| label: q2_part_b
percentile_aug <- augment(percentile_fit, new_data = movies_test)
percentile_aug |>
  count(percentile, .pred_class) |>
  group_by(percentile) |>
  mutate(p = n / sum(n))
```

# Part 2 - Credit Cards

The data for the second part of the take-home exam is on credit card
balances.

The variables in this dataset and their descriptions are as follows:

| Variable          | Description                                                            |
|-------------------|------------------------------------------------------------------------|
| `balance`         | Credit card balance in \$                                              |
| `income`          | Income in \$1,000                                                      |
| `student_status`  | Whether the individual is a student (`Student`) or not (`Not student`) |
| `marriage_status` | Whether the individual is a married (`Married`) or not (`Not married`) |
| `limit`           | Credit limit                                                           |

Assume that these data represent a random sample of American adults.

## Question 3: Model compare

**a.** The dataset is in the `data` folder of your repository, and it’s called
`credit.csv`. First, load the data with `read_csv()`, and assign it to
an object called `credit`.

```{r}
#| label: q3_part_a
credit <- read_csv("data/credit.csv")
```

**b.** Fit a model predicting `balance` from all other variables in the
dataset. This should be an additive model, i.e., use only main effects,
no interaction effects. Display a tidy output of the model.

```{r}
#| label: q3_part_b
balance_fit <- linear_reg() |>
  fit(balance ~ ., data = credit)

tidy(balance_fit)
```

**c.** Fit a model predicting `balance` from all other variables in the
dataset, **except for one of your choice**. This should be an additive
model, i.e., use only main effects, no interaction effects. Display a
tidy output of the model. Then, write the fitted equation of the model using proper statistical notation.

```{r}
#| label: q3_part_c
balance_fit_2 <- linear_reg() |>
  fit(balance ~ income + student_status + limit, data = credit)

tidy(balance_fit_2)
```

$$
\widehat{\text{balance}} = -429.51 -7.90 \times \text{income} + 419.68 \times \text {Student} + 0.27 \times \text{limit}
$$

**d.** Determine which model – the one from part (a) or the one from
part (b) – is the “better” model. Support your answer with an
appropriate summary statistic.

Based on a comparison of adjusted $R^2$ values, we would proceed with the model from part (b), which is an additive model with `income`, `student_status`, and `limit` as predictors.

```{r}
#| label: q3_part_d
glance(balance_fit)$adj.r.squared
glance(balance_fit_2)$adj.r.squared
```

**e.** For the model you chose in part (d), interpret the intercept and
**one** of the slopes in context of the data.

Intercept: For a non-student adult with a credit limit of \$0 and an income of \$0, we predict their credit card balance to be -\$429.51, on average.

Slope (Student): All else equal, we predict the credit card balance of a student adult to be \$419.68 larger than a non-student adult, on average.

## Question 4: Infer

What is the average difference in income between married and not married Americans?

**a.** Fit a simple linear regression model that estimates the difference in mean income between these two groups. Make sure you display the tidy output. Then, write the corresponding *population* model using proper statistical notation.

```{r}
#| label: q4_part_a
income_diff_fit <- linear_reg() |>
  fit(income ~ marriage_status, data = credit)

tidy(income_diff_fit)
```

$$
\text{income} = \beta_0 + \beta_1\times\text{Not married} + \epsilon
$$

**b.** Compute a 95% bootstrap interval for the slope of the regression line for predicting income (`income`) from marital status (`marriage_status`). In your code, use 1,000 bootstrap samples when simulating your bootstrap distribution.
Don't forget to set a seed!

In your narrative, first report your point estimate; then, provide an interpretation of the 95% confidence interval you obtain for the slope in the context of these data.

The point estimate is -2.81. We are 95% confident that, for non married adult, we predict their income to be between \$10.85 lower and \$4.99 higher than a married adult, on average.

```{r}
#| label: q4_part_b
set.seed(847)

observed_fit <- credit |>
  specify(income ~ marriage_status) |>
  fit()

observed_fit

boot_fits <- credit |>
  specify(income ~ marriage_status) |>
  generate(reps = 500, type = "bootstrap") |>
  fit()

ci_95 <- get_confidence_interval(boot_fits,
                                 point_estimate = observed_fit,
                                 level = .95,
                                 type = "percentile")

ci_95
```

**c.** Based on your answer to part (b), what would you expect your conclusion to be for a test of the following hypotheses at the 5% discernibility level, and why?

$$
\begin{aligned}
H_0: \beta_1 = 0 \\
H_A: \beta_1 \neq 0
\end{aligned}
$$
Because the 95% confidence interval contains 0, we would expect to fail to reject the null hypothesis above at the 5% discernibility level. 






