---
title: "Midterm 2 Study Guide"
execute:
  echo: false
---

::: callout-note
Answers are [here](/exams/midterm-2-answers.html).
:::

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(openintro)
library(scales)
library(palmerpenguins)
theme_set(theme_minimal(base_size = 12))
```

Midterm Exam 2 focuses on the modeling material that has been introduced since spring break. As such, you should be spending most of your time with these:

- [Lecture 15 (intro to models)](/slides/15-models.html)
    - [Application exercise 15 (ice cover)](/ae/ae-15-icecover-model.html)
- [Lecture 16 (simple linear regression)](/slides/16-simple-linear-regression.html)
    - [Application exercise 16 (penguins)](/ae/ae-16-modeling-penguins.html)
- [Lecture 17 (multiple linear regression)](/slides/17-multiple-linear-regression-1.html)
- [Lecture 18 (multiple linear regression)](/slides/18-multiple-linear-regression-2.html)
    - [Application exercise 18 (interest rates)](/ae/ae-18-modeling-loans.html)
- [Lecture 19 (logistic regression)](/slides/19-logistic-regression-1.html)
     - [Application exercise 19 (spam filter)](/ae/ae-19-spam-filter.html)
- [Lecture 20 (logistic regression)](/slides/20-logistic-regression-2.html)
     - [Application exercise 20 (forests)](/ae/ae-20-forest-classification.html)
- [Lecture 21 (logistic regression)](/slides/21-logistic-regression-3.html)
- [Lab 6](/lab/lab-6.html)
- [Lab 7](/lab/lab-7.html)
- [HW 5](/hw/hw-5.html)
- [HW 6](/hw/hw-6.html)

The exam consists of two parts:

- In-class portion
    - Wednesday April 8 at 11:45 AM - 1:00 PM in Griffith Theater;
    - Worth 80% of your Midterm 2 grade;
    - All multiple choice;
    - Your only resource is both sides of one 8.5" x 11" note sheet that you and only you created (written, typed, iPad, etc);
- Take-home portion
    - Released Thursday April 9 at 6:00 PM;
    - Due Sunday April 12 at 11:59 PM;
    - Worth 20% of your Midterm 2 grade;
    - Similar to a homework assignment, but shorter;
    - Completely open resource (open note, open internet, open ChatGPT, open everything), but citation policies apply, and collaboration between students is forbidden.
    
[Here](https://edstem.org/us/courses/90224/discussion/7626328) is our advice about how you should prioritize your studying. Below are 35 practice questions to get your wheels spinning.


::: callout-tip
## Try this

Attempt all of the problems below *without resources*. Make note of your pain points and blind spots along the way. That might give you an idea of the stuff you would benefit from if you had it on a cheat sheet.
:::

# Concepts and main ideas

### Question 1

Can you answer each of these in one or two sentences?

i. When we plot a line of best fit for a numerical predictor and numerical response, what does "best" mean?
i. What does the correlation coefficient between two numerical variables measure?
i. What does the $R^2$ of a linear regression measure?
i. What is the main lesson of examples like Anscombe's Quartet or the datasauRus dozen?
i. What is the difference between $Y=\beta_0+\beta_1 X + \varepsilon$ and $\hat{Y}=b_0+b_1 X$?
i. Under what circumstances is the intercept in a regression model a meaningful quantity?
i. If you run a "simple" linear regression model with a numerical response and a categorical predictor with three levels, what does the computer actually do to estimate that model?
i. Consider this bit of toy code: `linear_reg() |> fit(_BLANK_, df)`? What goes in the blank? What purpose does that argument serve, and how do you use it?
i. Let $\hat{y}^{(\text{old})}$ be the prediction at $x$ of a simple linear regression, and let $\hat{y}^{(\text{new})}$ be the prediction at $x+1$ of the same regression. How are $\hat{y}^{(\text{old})}$ and $\hat{y}^{(\text{new})}$ related? In other words, how does the prediction of the simple linear regression model change when you change the input from $x$ to $x+1$?
i. Let $\hat{o}^{(\text{old})}$ be the estimated odds at $x$ from a simple logistic regression model, and let $\hat{o}^{(\text{new})}$ be the estimated odds at $x+1$ of the same regression. How are $\hat{o}^{(\text{old})}$ and $\hat{o}^{(\text{new})}$ related? In other words, how does the prediction of the simple logistic regression model change when you change the input from $x$ to $x+1$?
i. Why is adjusted $R^2$ a better model selection criterion than unadjusted $R^2$?

# Visual understanding

### Question 2

Practice!

<https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>

### Questions 3 - 8

In what follows, let $y$ be a binary response, let $x$ be a numerical predictor, and let $z$ be a binary predictor. Match the S-curve to the model equation.

```{r}
#| fig-asp: 1
#| fig-width: 8
#| warning: false

set.seed(1234)

par(mfrow = c(3, 2))

n <- 100
x <- seq(-10, 10, length.out = n)
z <- rbinom(n, 1, 0.5)

# normal

b0 <- 0
b1 <- 0.5
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs3 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs3[1]
b1 <- coeffs3[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot 3")

# additive

b0 <- -3
b1 <- -1
b2 <- 2
y <- rbinom(n, 1, prob = exp(b0 + b1 * x + b2 * z) / (1 + exp(b0 + b1 * x + b2 * z)))

coeffs4 <- coefficients(glm(y ~ x + z, binomial(link = "logit")))
b0 <- coeffs4[1]
b1 <- coeffs4[2]
b2 <- coeffs4[3]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot 4")
curve(exp(b0 + b1 * x + b2) / (1 + exp(b0 + b1 * x + b2)), from = min(x), to = max(x), n = 1000, col = "blue", bty = "n", ylab = "", ylim = c(0, 1), add = TRUE)

# flipped

b0 <- 0
b1 <- -0.5
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs5 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs5[1]
b1 <- coeffs5[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot 5")

# shifted

b0 <- -5
b1 <- 1
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs6 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs6[1]
b1 <- coeffs6[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot 6")

# scrunched

b0 <- 20
b1 <- 4
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs7 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs7[1]
b1 <- coeffs7[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot 7")


# stretched

b0 <- 0.5
b1 <- 0.25
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs8 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs8[1]
b1 <- coeffs8[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot 8")

```

a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 22.9 + 4.54 \times x$
b. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.118 + 0.246 \times x$
c. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.798 + 0.486 \times x$
d. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = -6.035 + 1.194 \times x$
e. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = -2.26 - 0.99 \times x+1.26\times z$
f. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.079  -0.373 \times x$

### Questions 9 - 12

```{r}
#| echo: false
email <- email |> relocate(spam, exclaim_mess)
```


Recall the dataset on spam emails:

```{r}
#| echo: true
email
```

Here is a regression model trained on all of these data:

```{r}
#| echo: true
#| warning: false
#| message: false
log_fit <- logistic_reg() |>
  fit(spam ~ ., data = email)
```

The *in-sample* predictions of the model are here:

```{r}
#| echo: true
email_aug <- augment(log_fit, email)
email_aug
```

To see how well the model did, we can compute the classification error rates for different thresholds and summarize the results with the ROC curve:

```{r}
#| echo: true
email_roc <- roc_curve(email_aug, truth = spam, .pred_1, event_level = "second")
```


```{r}
#| echo: false

ggplot(email_roc, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  annotate("point", x = c(0.0, 0.0819, 0.229, 1.0), 
                    y = c(0.0, 0.531, 0.872, 1.0), 
                    size = 5) + 
  annotate("text", x = c(0.0 + 0.1, 0.2, 0.35, 1.0), 
                   y = c(0.0, 0.5, 0.872, 1.0 - 0.1), 
                   label = c("09", "10", "11", "12"), 
                   size = 5) +
  geom_abline(lty = 3) +
  coord_equal() + 
  labs(
    x = "1 - specificity",
    y = "sensitivity"
  )
```


Match the point on the ROC curve to the threshold that was used to compute it. 


a. 

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 0.25, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```

b. 

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 0.0, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```

c. 

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 1.00, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```

d.

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 0.1, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```


# WANANA doctor

What's wrong with the interpretations below? Diagnose the problem and fix it if possible.

```{r}
q_no <- 12

q_no <- q_no + 1
```

## Question `{r} q_no`

Some bozo ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(bill_length_mm ~ flipper_length_mm, penguins) |>
  tidy()
```

> If the flipper length of a penguin increases by 1mm, then the bill length will increase by 0.25mm.

```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some schmuck ran this regression and wrote the following interpretation:

```{r}
#| echo: true
logistic_reg() |>
  fit(spam ~ num_char, email) |>
  tidy()
```

> If the number of characters in an email increases by 1000, then the estimated probability of that email being spam is lower by -0.062, on average.

FYI: from the documentation (`?email`), `num_char` is "The number of characters in the email, in thousands."

```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some putz ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, penguins) |> 
  tidy()
```

> If flipper length and bill length both increase by 1mm, then we predict that body mass is lower by 48.14 grams, on average.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some blockhead ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm * species, penguins) |>
  tidy()
```

> If the flipper length of a Chinstrap penguin increases by 1mm, then we predict that the body mass is higher by 1.74 grams on average.

```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some nitwit ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(hp ~ mpg, mtcars) |>
  glance()
```

> About 60% of the observed variation in miles/gallon is explained by the model.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some chooch ran this regression and wrote the following interpretation:

```{r}
#| echo: true
logistic_reg() |>
  fit(spam ~ exclaim_mess + viagra + dollar, email) |>
  tidy()
```

> If an email contains no exclamation marks, no dollar signs, and no mentions of viagra, then the model predicts that the probability of it being spam is $e^{-2.218}\approx 0.10886$.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some nincompoop ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(bill_length_mm ~ bill_depth_mm + sex, penguins) |>
  tidy()
```

> If a male penguin has a bill depth of 0mm (sad!), then we expect its bill length to be 60.99mm on average.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some silly goose ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, penguins) |>
  glance()
```

> The correlation between body mass, flipper length, and bill length is about $\sqrt{0.76}\approx 0.87$.



# Data analysis

```{r}
#| label: blizzard-data-prep
blizzard_salary <- blizzard_salary |>
  mutate(
    annual_salary = case_when(
      salary_type == "week" ~ current_salary * 52,
      salary_type == "hour" ~ current_salary * 40 * 52,
      TRUE ~ current_salary
    ),
    performance_rating = if_else(
      performance_rating == "Developing",
      "Poor",
      performance_rating
    )
  ) |>
  filter(salary_type != "week") |>
  mutate(
    salary_type = if_else(salary_type == "hour", "Hourly", "Salaried")
  ) |>
  filter(!is.na(annual_salary)) |>
  select(percent_incr, salary_type, annual_salary, performance_rating)
```

## Blizzard salaries

In 2020, employees of Blizzard Entertainment circulated a spreadsheet to anonymously share salaries and recent pay increases amidst rising tension in the video game industry over wage disparities and executive compensation.
(Source: [Blizzard Workers Share Salaries in Revolt Over Pay](https://www.bloomberg.com/news/articles/2020-08-03/blizzard-workers-share-salaries-in-revolt-over-wage-disparities))

The name of the data frame used for this analysis is `blizzard_salary` and the variables are:

-   `percent_incr`: Raise given in July 2020, as percent increase with values ranging from 1 (1% increase) to 21.5 (21.5% increase)

-   `salary_type`: Type of salary, with levels `Hourly` and `Salaried`

-   `annual_salary`: Annual salary, in USD, with values ranging from \$50,939 to \$216,856.

-   `performance_rating`: Most recent review performance rating, with levels `Poor`, `Successful`, `High`, and `Top`.
    The `Poor` level is the lowest rating and the `Top` level is the highest rating.

The first ten rows of `blizzard_salary` are shown below:

```{r}
blizzard_salary |>
  select(percent_incr, salary_type, annual_salary, performance_rating)
```

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

You fit a model for predicting raises (`percent_incr`) from salaries (`annual_salary`). We'll call this model `raise_1_fit`. A tidy output of the model is shown below.

```{r}
#| label: raise-salary-fit
raise_1_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary, data = blizzard_salary)

tidy(raise_1_fit)
```

Which of the following is the best interpretation of the slope coefficient?

a.  For every additional \$1,000 of annual salary, the model predicts the raise to be higher, on average, by 1.55%.
b.  For every additional \$1,000 of annual salary, the raise goes up by 0.0155%.
c.  For every additional \$1,000 of annual salary, the model predicts the raise to be higher, on average, by 0.0155%.
d.  For every additional \$1,000 of annual salary, the model predicts the raise to be higher, on average, by 1.87%.

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

You then fit a model for predicting raises (`percent_incr`) from salaries (`annual_salary`) and performance ratings (`performance_rating`). We'll call this model `raise_2_fit`. Which of the following is definitely true based on the information you have so far?

a.  Intercept of `raise_2_fit` is higher than intercept of `raise_1_fit`.
b.  Slope of `raise_2_fit` is higher than RMSE of `raise_1_fit`.
c.  Adjusted $R^2$ of `raise_2_fit` is higher than adjusted $R^2$ of `raise_1_fit`.
d.  $R^2$ of `raise_2_fit` is higher $R^2$ of `raise_1_fit`.


```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

The tidy model output for the `raise_2_fit` model you fit is shown below.

```{r}
#| label: raise-salary-rating-fit
raise_2_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary + performance_rating, data = blizzard_salary)

tidy(raise_2_fit)
```

When your teammate sees this model output, they remark 

> "The coefficient for `performance_ratingSuccessful` is negative. That's weird. I guess it means that people who get successful performance ratings get lower raises." 

How would you respond to your teammate?

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Ultimately, your teammate decides they don't like the negative slope coefficients in the model output you created (not that there's anything wrong with negative slope coefficients!), does something else, and comes up with the following model output.

```{r}
blizzard_salary <- blizzard_salary |>
  mutate(
    performance_rating = fct_relevel(
      performance_rating,
      "Poor",
      "Successful",
      "High",
      "Top"
    )
  )

raise_2_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary + performance_rating, data = blizzard_salary)

tidy(raise_2_fit)
```

Unfortunately they didn't write their code in a Quarto document, instead just wrote some code in the Console and then lost track of their work. They remember using the `fct_relevel()` function and doing something like the following:

```{r}
#| eval: false
#| echo: true
blizzard_salary <- blizzard_salary |>
  mutate(performance_rating = fct_relevel(performance_rating, ___))
```

What should they put in the blanks to get the same model output as above?

a.  "Poor", "Successful", "High", "Top"
b.  "Successful", "High", "Top"
c.  "Top", "High", "Successful", "Poor"
d.  Poor, Successful, High, Top

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Suppose we fit a model to predict `percent_incr` from `annual_salary` and `salary_type`. A tidy output of the model is shown below.

```{r}
#| label: raise-salary-type-fit
raise_3_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary + salary_type, data = blizzard_salary)

tidy(raise_3_fit)
```

Which of the following visualizations represent this model? Explain your reasoning.

```{r}
#| label: fig-raise-salary-type
#| warning: false
#| layout-ncol: 2
#| fig-cap: |
#|   Visualizations of the relationship between percent increase, annual
#|   salary, and salary type
#| fig-subcap:
#|   - Option 1
#|   - Option 2
#|   - Option 3
#|   - Option 4
ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(aes(shape = salary_type), alpha = 0.5, size = 2) +
  geom_smooth(
    aes(linetype = salary_type),
    method = "lm",
    se = FALSE,
    fullrange = TRUE,
    linewidth = 1.5
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase",
    color = "Salary type",
    linetype = "Salary type",
    shape = "Salary type"
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  theme(legend.position = "top")

ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(aes(shape = salary_type), alpha = 0.5, size = 2) +
  geom_smooth(
    aes(linetype = salary_type),
    se = FALSE,
    fullrange = TRUE,
    linewidth = 1.5
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase"
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  theme(legend.position = "top")

ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(
    aes(shape = salary_type),
    alpha = 0.5,
    size = 2,
    show.legend = FALSE
  ) +
  geom_abline(
    intercept = 1.24,
    slope = 0.0000137,
    color = "#E87d72",
    linewidth = 1.5,
    linetype = "solid"
  ) +
  geom_abline(
    intercept = 1.24 + 0.913,
    slope = 0.0000137,
    color = "#56bcc2",
    linewidth = 1.5,
    linetype = "dashed"
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase",
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1))

ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(alpha = 0.5, size = 2, show.legend = FALSE) +
  geom_abline(
    intercept = 1.24,
    slope = 0.0000137,
    color = "#56bcc2",
    linewidth = 1.5,
    linetype = "dashed"
  ) +
  geom_abline(
    intercept = 1.24 + 0.913,
    slope = 0.0000137,
    color = "#E87d72",
    linewidth = 1.5
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase",
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1))
```



```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Suppose you now fit a model to predict the natural log of percent increase, `log(percent_incr)`, from performance rating. The model is called `raise_4_fit`.

```{r}
raise_4_fit <- linear_reg() |>
  fit(log(percent_incr + 0.0001) ~ performance_rating, data = blizzard_salary)
```

You're provided the following:

```{r}
#| echo: true
tidy(raise_4_fit) |>
  select(term, estimate) |>
  mutate(exp_estimate = exp(estimate))
```

Based on this, which of the following is true?

a\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by 10.25% compared to the employees with Poor performance rating.

b\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by 6.93% compared to the employees with Poor performance rating.

c\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by a factor of 1025 compared to the employees with Poor performance rating.

d\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by a factor of 6.93 compared to the employees with Poor performance rating.



## Movies

The data for this part comes from the Internet Movie Database (IMDB). Specifically, the data are a random sample of movies released between 1980 and 2020.

```{r}
#| label: load-data
#| message: false
movies <- read_csv("data/movies.csv")
```

The name of the data frame used for this analysis is `movies`, and it contains the variables shown in @tbl-data-dictionary.

\setcounter{table}{0}

| Variable          | Description                                                             |
|----------------------|--------------------------------------------------|
| `name`            | name of the movie                                                       |
| `rating`          | rating of the movie (R, PG, etc.)                                       |
| `genre`           | main genre of the movie.                                                |
| `runtime`         | duration of the movie                                                   |
| `year`            | year of release                                                         |
| `release_date`    | release date (YYYY-MM-DD)                                               |
| `release_country` | release country                                                         |
| `score`           | IMDB user rating                                                        |
| `votes`           | number of user votes                                                    |
| `director`        | the director                                                            |
| `writer`          | writer of the movie                                                     |
| `star`            | main actor/actress                                                      |
| `country`         | country of origin                                                       |
| `budget`          | the budget of a movie (some movies don't have this, so it appears as 0) |
| `gross`           | revenue of the movie                                                    |
| `company`         | the production company                                                  |

: Data dictionary for `movies` {#tbl-data-dictionary tbl-colwidths="\[20,80\]"}

The first thirty rows of the `movies` data frame are shown in @tbl-data, with variable types suppressed (since we'll ask about them later).


```{r}
#| echo: false
movies_to_mark <- c("Blue City", "Rang De Basanti", "Winter Sleep")

movies <- movies |>
  mutate(
    mark = if_else(name %in% movies_to_mark, TRUE, FALSE),
    rating = case_when(
      rating == "TV-PG" ~ "PG",
      rating == "Unrated" ~ "Not Rated",
      is.na(rating) ~ "Not Rated",
      .default = rating
    ),
    rating = fct_relevel(rating, "G", "PG", "PG-13", "R", "NC-17", "Not Rated")
  ) |>
  arrange(desc(mark)) |>
  relocate(
    name,
    score,
    runtime,
    genre,
    rating,
    release_country,
    release_date,
    budget,
    gross,
    votes,
    year,
    director,
    writer,
    star,
    company,
    country
  )
```

```{=tex}
\begin{landscape}
```

::: {#tbl-data}
First 30 rows of `movies`, with variable types suppressed.

```{r}
#| echo: false
options(
  dplyr.print_min = 30,
  pillar.min_chars = 13,
  pillar.width = 80,
  pillar.sigfig = 6
)

format(movies |> select(!c(mark)))[-3L] |>
  str_remove_all(" <.*?>") |>
  cat(sep = "\n")

options(
  dplyr.print_min = 10,
  pillar.min_chars = 8,
  pillar.width = 80,
  pillar.sigfig = 3
)
```

:::



### Score vs. runtime

In this part, we fit a model predicting `score` from `runtime` and name it `score_runtime_fit`.

```{r}
#| echo: true
score_runtime_fit <- linear_reg() |>
  fit(score ~ runtime, data = movies)
```


@fig-score-runtime visualizes the relationship between `score` and `runtime` as well as the linear model for predicting `score` from `runtime`. The first three movies in @tbl-data are labeled in the visualization as well. Answer all questions in this part based on @fig-score-runtime.

```{r}
#| label: fig-score-runtime-prep
#| echo: false
movies <- movies |>
  mutate(
    runtime = parse_number(runtime),
    shape1 = case_when(
      name == movies_to_mark[1] ~ "circle",
      name == movies_to_mark[2] ~ "square",
      name == movies_to_mark[3] ~ "triangle",
      .default = "circle"
    ),
    shape2 = case_when(
      name == movies_to_mark[1] ~ "circle open",
      name == movies_to_mark[2] ~ "square open",
      name == movies_to_mark[3] ~ "triangle open",
      .default = NA
    ),
    color = if_else(mark, "black", "gray50"),
    alpha = if_else(mark, 1, 0.5)
  )
```

```{r}
#| label: fig-score-runtime
#| fig-cap: Scatterplot of `score` vs. `runtime` for `movies`.
#| echo: false
#| message: false
#| fig-asp: 0.5
#| fig-width: 6.5
ggplot(movies, aes(x = runtime, y = score)) +
  geom_point(aes(shape = shape1, color = color, alpha = alpha)) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_shape_identity() +
  scale_alpha_identity() +
  scale_color_identity() +
  geom_point(
    data = movies |> filter(mark),
    aes(shape = shape2),
    size = 3
  ) +
  geom_label(
    data = movies |> filter(mark),
    aes(label = name),
    size = 3,
    nudge_x = 1,
    nudge_y = c(-0.4, -0.4, 0.4)
  ) +
  scale_x_continuous(breaks = seq(80, 240, 20))
```


```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Partial code for producing @fig-score-runtime is given below. Which of the following goes in the blank on Line 2? **Select all that apply.**


```{r}
#| eval: false
#| echo: true
#| code-line-numbers: true
movies |>
  mutate(runtime = ___) |>
  ggplot(aes(x = runtime, y = score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE)
# additional code for annotating Blue City on the plot
```

a.  `grepl(" mins", runtime)`

b.  `grep(" mins", runtime)`

c.  `str_remove(runtime, " mins")`

d.  `as.numeric(str_remove(runtime, " mins"))`

e.  `na.rm(runtime)`

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Based on this model, order the three labeled movies in @fig-score-runtime in decreasing order of the magnitude (absolute value) of their residuals.

a.  Winter Sleep \> Rang De Basanti \> Blue City

b.  Winter Sleep \> Blue City \> Rang De Basanti

c.  Rang De Basanti \> Winter Sleep \> Blue City

d.  Blue City \> Winter Sleep \> Rang De Basanti

e.  Blue City \> Rang De Basanti \> Winter Sleep



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

```{r}
#| include: false
score_runtime_rsq <- round(glance(score_runtime_fit)$r.squared * 100, 0)
```

The R-squared for the model visualized in @fig-score-runtime is `{r} score_runtime_rsq`%. Which of the following is the [**best**]{.underline} interpretation of this value?

a.  `{r} score_runtime_rsq`% of the variability in movie runtimes is explained by their scores.

b.  `{r} score_runtime_rsq`% of the variability in movie scores is explained by their runtime.

c.  The model accurately predicts scores of `{r} score_runtime_rsq`% of the movies in this sample.

d.  The model accurately predicts scores of `{r} score_runtime_rsq`% of all movies.

e.  The correlation between scores and runtimes of movies is `{r} score_runtime_rsq / 100`.

### Score vs. runtime or year

The visualizations below show the relationship between `score` and `runtime` as well as `score` and `year`, respectively. Additionally, the lines of best fit are overlaid on the visualizations.

```{r}
#| layout-ncol: 2
#| echo: false
#| message: false
ggplot(movies, aes(x = runtime, y = score)) +
  geom_point(alpha = 0.5, color = "gray40") +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 1.2)

ggplot(movies, aes(x = year, y = score)) +
  geom_point(alpha = 0.5, color = "gray40") +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 1.2)
```

The correlation coefficients of these relationships are calculated below, though some of the code and the output are missing. Answer all questions in this part based on the code and output shown below.

```{r}
#| include: false
movies |>
  summarize(
    r_score_runtime = cor(runtime, score),
    r_score_year = cor(year, score)
  )
```

```{r}
#| eval: false
#| echo: true
movies |>
  __blank_1__(
    r_score_runtime = cor(runtime, score),
    r_score_year = cor(year, score)
  )
```

```         
# A tibble: 1 × 2
  r_score_runtime r_score_year
            <dbl>        <dbl>
1           0.434. __blank_2__       
```

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Which of the following goes in `__blank_1__`?

a.  `summarize`

b.  `mutate`

c.  `group_by`

d.  `arrange`

e.  `filter`

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

What can we say about the value that goes in `__blank_2__`?

a.  `NA`

b.  A value between 0 and 0.434.

c.  A value between 0.434 and 1.

d.  A value between 0 and -0.434.

e.  A value between -1 and -0.434.


### Score vs. runtime and rating

In this part, we fit a model predicting `score` from `runtime` and `rating` (categorized as G, PG, PG-13, R, NC-17, and Not Rated), and name it `score_runtime_rating_fit`.

```{r}
score_runtime_rating_fit <- linear_reg() |>
  fit(score ~ runtime + rating, data = movies)
```

The model output for `score_runtime_rating_fit` is shown in @tbl-score-runtime-rating-tidy. Answer all questions in this part based on @tbl-score-runtime-rating-tidy.

```{r}
#| label: tbl-score-runtime-rating-tidy
#| tbl-cap: Regression output for `score_runtime_rating_fit`.
#| echo: false
tidy(score_runtime_rating_fit) |>
  knitr::kable(digits = 3)
```



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Which of the following is [**TRUE**]{.underline} about the intercept of `score_runtime_rating_fit`? **Select all that are true.**

a.  Keeping runtime constant, G-rated movies are predicted to score, on average, 4.525 points.

b.  Keeping runtime constant, movies without a rating are predicted to score, on average, 4.525 points.

c.  Movies without a rating that are 0 minutes in length are predicted to score, on average, 4.525 points.

d.  All else held constant, movies that are 0 minutes in length are predicted to score, on average, 4.525 points.

e.  G-rated movies that are 0 minutes in length are predicted to score, on average, 4.525 points.



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Which of the following is the [**best**]{.underline} interpretation of the slope of `runtime` in `score_runtime_rating_fit`?

a.  All else held constant, as runtime increases by 1 minute, the score of the movie increases by 0.021 points.

b.  For G-rated movies, all else held constant, as runtime increases by 1 minute, we predict the score of the movie to be higher by 0.021 points on average.

c.  All else held constant, for each additional minute of runtime, movie scores will be higher by 0.021 points on average.

d.  G-rated movies that are 0 minutes in length are predicted to score 0.021 points on average.

e.  For each higher level of rating, the movie scores go up by 0.021 points on average.

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Fill in the blank:

> R-squared for `score_runtime_rating_fit` (the model predicting `score` from `runtime` and `rating`) \_\_\_\_\_\_\_\_\_ the R-squared the model `score_runtime_fit` (for predicting `score` from `runtime` alone).

a.  is less than

b.  is equal to

c.  is greater than

d.  cannot be compared (based on the information provided) to

e.  is both greater than and less than



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

The model `score_runtime_rating_fit` (the model predicting `score` from `runtime` and `rating`) can be visualized as parallel lines for each level of `rating`. Which of the following is the equation of the line for R-rated movies?

a.  $\widehat{score} = (4.525 - 0.257) + 0.021 \times runtime$

b.  $score = (4.525 - 0.257) + 0.021 \times runtime$

c.  $\widehat{score} = 4.525 + (0.021 - 0.257) \times runtime$

d.  $score = 4.525 + (0.021 - 0.257) \times runtime$

e.  $\widehat{score} = (4.525 + 0.021) - 0.257 \times runtime$
