---
title: "Milestone 4"
date: "2025-06-16"
format: revealjs
auto-stretch: false
editor_options: 
  chunk_output_type: console
---

## Announcements

-   Commit and pushes changes to AE

-   TEAMMATES peer evaluation will open today at noon and close Wednesday at 11:59pm

    -   Let Marie/Mary know if you are not getting emails from TEAMMATES (check spam folder)

-   Today

    -   Tips to help make project beautiful

    -   Work on project

    -   Review one other team's project

## Code cell options {.smaller}

-   We will examine the `echo`, `warning`, `eval`, `fig-width`, and `fig-asp` code chunk options using the following code:

```{r}
#| message: false

library(tidyverse)
library(palmerpenguins)
```

```{r}
#| eval: false
#| echo: true
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```

-   To set code chunk options:

    -   `#| option: value` -- The spaces must be included! Without first space it's just a comment.

    -   Must be at the top of the code chunk or within the YAML at the top of the document (to include the same option for all code chunks)

-   Try the various options, render the document, and see what happens


## Code cell options -- echo {.smaller}

::::: columns
::: {.column width="50%"}
```{r}
#| echo: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::

::: {.column width="50%"}
```{r}
#| echo: true

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::
:::::

::: incremental
-   echo: Includes the code chunk in the ouptut.

-   We want this for labs to show our work. We do NOT want this for the project

-   To include this for the main project file (index.qmd), within yaml at top of document, change `echo: true` to `echo: false`. Then `echo` is set to `false` for ALL code chunks
:::

## Code cell options -- warning {.smaller}

::::: columns
::: {.column width="50%"}
```{r}
#| warning: false
#| echo: true

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::

::: {.column width="50%"}
```{r}
#| warning: true
#| echo: true

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::
:::::

::: incremental
-   warning: Includes warning messages in the output.
:::

## Code cell options -- eval {.smaller}

::::: columns
::: {.column width="50%"}
```{r}
#| eval: false
#| echo: true

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::

::: {.column width="50%"}
```{r}
#| eval: true
#| echo: true

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::
:::::

::: incremental
-   eval: Evaluates, or runs, the code.

-   Can be helpful if you are unable to fix an error before the deadline.
:::

## Code cell options -- fig-width {.smaller}

::::: columns
::: {.column width="50%"}
```{r}
#| fig-width: 1
#| echo: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::

::: {.column width="50%"}
```{r}
#| fig-width: 4
#| echo: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::
:::::

-   fig-width: controls figure width (is a value between 1 and 10)

-   `fig-width: 1` on the left, `fig-width: 4` on the right

## Code cell options -- fig-asp {.smaller}

::::: columns
::: {.column width="33%"}
```{r}
#| fig-asp: 0.1
#| echo: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::

::: {.column width="33%"}
```{r}
#| fig-asp: 0.5
#| echo: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::

::: {.column width="33%"}
```{r}
#| fig-asp: 1
#| echo: false

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
:::
:::::

-   fig-asp: controls figure aspect ratio (is a value between 0 and 1)

-   `fig-asp: 0.1` on the left, `fig-asp: 0.5` in the middle, `fig-asp: 1` on the right

## Tables {.smaller}

-   Can create pretty tables using `kable` (uses knitr library)

```{r}
library(knitr)
```

::::: columns
::: {.column width="70%"}
```{r}
#| echo: true



mtcars |>
  group_by(cyl) |>
  summarize(
    avg_mpg = mean(mpg),
    avg_hp = mean(hp),
    count = n()
  ) |>
  kable(digits = 2, 
        caption = "Average MPG and HP by Cylinder")
```
:::

::: {.column width="30%"}
```{r}
#| echo: true
mtcars |>
  group_by(cyl) |>
  summarize(
    avg_mpg = mean(mpg),
    avg_hp = mean(hp),
    count = n()
  ) 
```
:::
:::::

